---
title: "Determine co-localization in scRNA using seurat intergration"
output: html_notebook
---

## Summary of script  
1. Read seurat object
2. determine patient information is uniformly available 
3. If not 2, then read gse meta data to determine the infomation.
4. Summarize the annotation, and gd-T alleles
5. If Trgd  is absent use the blood atlas as marker 
6. Try w/o annotation

```{r}
files = list.files(pattern="*.res.rds", full.names = T, path = "~/liulab_home/data/links/data/", recursive = T)
nonicb.obj1 = list.files(path="/liulab/zzeng/Tigger/static/data/scNonICB/", pattern = "*.rds",recursive = T, full.names = T)

```

## Perform linear mixed model 

```{r}
library(parallel)
library(nlme)
library(Matrix)
extract_nmle_table <- function (m1){
    mod = summary(m1)
    beta <- m1$coefficients$fixed[2] #$fixed is not needed
    se <- m1$varFix[2]
    t <- beta/se
    p<- anova(m1)[2,4]
    table=data.frame(cbind(beta,se,t,p))
    return(table)
}
            
eval.nlme = function(data.dt){
    tryCatch(
        {
            data.dt = data.dt[!is.na(y)]
            data.dt$x %<>% avinash::znorm()
            data.dt$y %<>% avinash::znorm()
            m1 <- lme(y~ x, random=~1|dataset, data=data.dt)
            extract_nmle_table(m1)
        },error=  function(e) rep(NA,4))
}

perform.lmer.coinfiltration =  function(dt.x, dt.y, label = NULL) {
  dt.x$y = dt.y[match(dt.x$patient.new, patient.new)]$frac 
  dt.x[,x:=frac]
  out = dt.x[,.(list(eval.nlme(.SD))), by=clust]
  do.call(rbind, out$V1) %>% data.table %>% 
    .[,clust:=out$clust]%>%
    .[order(p)] %>% 
    .[,label:=label]
}
calcualte.immune.fraction.davis = function(meta.data){
  if(is.null(meta.data$patient))  return(NULL)
  meta.data %<>% as.data.table
  col.sel = "assign.level3_anno" 
  meta.data[,total.count:=.N, by=patient]
  sapply( col.sel, function(tt) meta.data[,.(frac=.N/unique(total.count)),by=c("patient", tt)], 
                USE.NAMES=T, simplify = F)
}


create.psuedobulk <- function(exp.mat, labels) {
  dt1 = data.table(V1=labels)
  one.hot = mltools::one_hot(dt1) %>% as.matrix
  one.hot %>% colMeans %>% 
    sweep(one.hot,., MARGIN = 2, FUN = "/")
  out = exp.mat %*% one.hot
  out %>% set_colnames(gsub("V1_", colnames(one.hot), replacement = ""))
}
```

## Select genes to store

```{r}

get.synonym = function(gene){
  geneSynonym::humanSyno(gene,  caseSensitive = F) %>% 
  unlist %>% toupper 
}


tgd.specific.genes = c("TMEM132C", "GALNT18", "FBXO39", "NFASC", "HNF4G","KRT84", "CYP2C8", "LIM2", "TFCP2L1", 
	"AC012531.3", "KLKB1", "MFAP2")
gdt.blood.marker = fread("~/liulab_home/data/single_cell/markers/./blood_cell_category_rna_gdT-cell_Cell.tsv")
gdt.secreted = c("GM-CSF", "GZMB", "IFNG", "IL4", "IL5", "IL17A", "PRF1", "TNFA")
gdt.secreted.pruned = c("CSF2", "GZMB", "IFNG", "IL4", "IL5", "IL17A", "PRF1", "TNF")
gdt.surface=c("CD5", "CD16", "CD27", "CD28", "CD45RA", "CD56", "CD57", "CD62L", "CD69", "CD70H", "CD107a", "NKG2D")
gdt.surface=c("CD5", "FCGR3A", "FCGR3B", "CD27", "CD28", "CD3G", "NCAM1", "B3GAT1", "SELL", "CD69", "CD70", "LAMP1", "KLRK1")

all.genes = c(tgd.specific.genes, gdt.blood.marker, gdt.secreted, gdt.secreted.pruned, gdt.surface, gdt.surface) %>% unique()
all.genes = all.genes %>% get.synonym() 
```


Get transferred label of datasets 
```{r}
get.dataset.cluster <- function(obj.file, transfer.label.files) {
  obj.name = basename(obj.file)
  cluster = NULL 
 transfer.label.file =  grep(obj.name, transfer.label.files, value=T)
  if(length(transfer.label.file) > 0) {
    readRDS(transfer.label.file)$predicted.id  -> cluster
  }
 cluster 
 }

transfer.label.files = list.files("~/liulab_home/data/single_cell/transfer.labels.datasets", full.names = T)
dataset.label = lapply( nonicb.obj1, get.dataset.cluster, transfer.label.files =transfer.label.files )
icb.seurat_cluster = readRDS("/homes6/asahu/liulab_home/data/single_cell//all.icb.integrated/seurat_cluster.rds")
```


## Read annotation from TISCH 
```{r}
library(Seurat)
exps = mclapply(files, function(file){
  sco = readRDS(file)
  label = file %>% basename %>% gsub("_res.rds", "", .)
  genes.curr = rownames(sco$RNA)
  tr.alleles = grep("^TR[ABDG][VDJC]", genes.curr, value=T)
  tr.alleles = c(tr.alleles, "CD8A", "CD3E", "CD3G") 
  genes.sel = which(genes.curr %in% c( tr.alleles, all.genes))
  list(label =label, meta.data = sco$RNA@meta.data, exp = sco$RNA@assays$RNA@data[genes.sel,])
}, mc.cores = 48
)
names.exps = sapply(exps, function(tt) tt$label)
names(exps) = names.exps

```


## Mean of alleles 
Using ab-T cells as control comapre gdT interaction. I am using all mean of alleles of gamma delta

```{r}

calculate.immune.fraction = function(meta.data){
  if(is.null(meta.data$patient))  return(NULL)
  meta.data %<>% as.data.table
  col.sel = grep("assign", colnames(meta.data), value  =T )
  meta.data[,total.count:=.N, by=patient]
  sapply( col.sel, function(tt) meta.data[,.(frac=.N/unique(total.count)),by=c("patient", tt)], 
                USE.NAMES=T, simplify = F)
}

#' Calculate gdT score based on different gene signature
#'
#' @param object list with meta.data and exp  
#' @param genes list of genes (default:NULL)
#' @param grep.genes pattern of grepping genes (default:NULL)
#' @param genes.weight list of genes weight (default:NULL)
#'
#' @return
#' @export
#'
#' @examples
calculate.gdT.score = function(object, genes = NULL, grep.genes = NULL, genes.weight =NULL){
  exp.curr = object$exp
  patient = object$meta.data$patient
  label = object$label
  if(is.null(patient)) return(NULL) 
  if(!is.null(grep.genes)) {
    genes %<>% c(., grep(grep.genes, rownames(exp.curr), value=T) )
  }
   genes %<>% intersect(rownames(exp.curr))
  if(length(genes)==0) return(NULL)
  weights.curr = matrix(0, nrow=1, ncol=nrow(exp.curr)) %>% 
    set_colnames(rownames(exp.curr))
  weights.curr[,genes] =1
  if(!is.null(genes.weight)){
    genes.weight = genes.weight[intersect(names(genes.weight), rownames(exp.curr))]
    weights.curr[,names(genes.weight)] = unlist(genes.weight)
  }
  weights.curr = weights.curr/sum(weights.curr)
   cell.score = weights.curr %*% exp.curr %>% as.matrix %>% c()
  data.table(cell.score=cell.score, patient=patient) %>% 
    .[,.(patient.score=mean(cell.score)), by=patient] %>% 
    .[,dataset:=label] %>% 
    .[,patient.new:=paste(patient, label, sep = ":")]
}
```


```{r}
abT.scores =  mclapply(exps, function(tt) calculate.gdT.score(tt, grep.genes = "^TR[AB]"), mc.cores = 48)
gdT.scores =  mclapply(exps, function(tt) calculate.gdT.score(tt, grep.genes = "^TR[GD]"), mc.cores = 48)
abT.scores = do.call(rbind, abT.scores)
gdT.scores = do.call(rbind, gdT.scores)
cor.test(temp1$patient.score, temp2$patient.score)
aa = merge(temp1, temp2, by="patient")

immune.fracs = mclapply(exps, function(tt) calcualte.immune.fraction(tt$meta.data), mc.cores = 48)
## find unique annotation mechanism 
annot.mechanism = sapply(immune.fracs, names) %>% unlist %>% unique 
agg.immune.fracs = list()
for (ii in seq_along(immune.fracs)) {
  immune.frac = immune.fracs[[ii]]
  label.curr = exps[[ii]]$label 
  for (tt in seq_along(immune.frac)) {
    annot.curr = names(immune.frac)[[tt]]
    setnames(immune.frac[[tt]], 2, "clust")
    immune.frac[[tt]][,dataset:=label.curr][,annot:=annot.curr][,patient.new:=paste(patient, label.curr, sep = ":")]
    agg.immune.fracs[[annot.curr]] %<>%  rbind(., immune.frac[[tt]])
  }
}
# for (ii in seq_along(agg.immune.fracs)) {
#    agg.immune.fracs[[ii]] %<>% dcast(., patient.new  ~ clust, value.var = "frac")
# }

```


```{r}
agg.immune.fracs.old = agg.immune.fracs
agg.immune.fracs.new = agg.immune.fracs
agg.immune.fracs.new$assign.score = NULL
abT.coinfiltration = lapply(names(agg.immune.fracs.new),  function(tt) perform.lmer.coinfiltration(dt.x=agg.immune.fracs[[tt]], dt.y=abT.scores[,frac:=patient.score], label=tt))
abT.coinfiltration %<>% do.call(rbind,.) %>% 
  .[order(p)]
gdT.coinfiltration = lapply(names(agg.immune.fracs.new),  function(tt) perform.lmer.coinfiltration(dt.x=agg.immune.fracs[[tt]], dt.y=gdT.scores[,frac:=patient.score], label=tt))
gdT.coinfiltration %<>% do.call(rbind,.) %>% 
  .[order(p)]

# aa = lapply(agg.immune.fracs, function(tt) any(is.numeric(tt$clust)))

coinfiltration.dt = merge(abT.coinfiltration %>% as.data.frame(), gdT.coinfiltration %>% as.data.frame(), by=c("clust", "label"), suffixes = c(".ab",".gd")) %>% as.data.table()
get.sign.log = function(tt){
  sign(tt)*log(abs(tt))
} 
coinfiltration.dt$t.ab %<>% get.sign.log
coinfiltration.dt$t.gd %<>% get.sign.log
coinfiltration.dt[,logp.ab:=-sign(t.ab)*log10(p.ab)] 
coinfiltration.dt[,logp.gd:=-sign(t.gd)*log10(p.gd)] 
coinfiltration.dt.sub = coinfiltration.dt[which((abs(logp.ab) > 4) | (abs(logp.gd) > 2.5))]
p.com = ggplot(data=coinfiltration.dt, aes(x=logp.ab, y=logp.gd))  + 
  geom_point(alpha=0.5) +
  theme_minimal(base_size = 12) + theme(legend.position = "bottom") +
  geom_text_repel(
    data = coinfiltration.dt.sub,
    aes(x=logp.ab, y=logp.gd, label = clust),    
    size = 2,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  ) + 
  ylab("gamma-delta association")  + 
  xlab("alpha-beta association") + 
  theme_bw()
ggsave("/liulab/asahu/projects/gdt/results/.figs/coinfiltration.gd.ab.pdf", p.com)


change.dt =coinfiltration.dt%>% 
  .[,clust:=as.character(clust)]
change.dt[,clust:=ifelse(clust=="Immune cells", "Memory_B_cells", change.dt$clust)]
change.dt[,clust:=ifelse(clust=="Microglia", "Fibroblasts", change.dt$clust)]
p = EnhancedVolcano::EnhancedVolcano(change.dt,
	lab = change.dt$clust,
	x = 'beta.gd',
	y = 'p.gd',
	pCutoff = 1E-2,
	FCcutoff = .01,
	labSize = 4.0,
	legend=c('NS','Log (base 2) fold-change','Adj.P value',
		'Adj.P value & Log (base 2) fold-change'),
	legendPosition = 'right',
	# legendLabSize = 8,
	# legendIconSize = 4.0,
	drawConnectors = TRUE,
	widthConnectors = 0.2,
	colAlpha = 0.8,
	colConnectors = 'grey30'
	)


```
## TRDV1 and TRDV2
```{r}
get.coinfiltration = function(genes = NULL, grep.genes=NULL, agg.immune.fracs = agg.immune.fracs){
  pattern.scores =  mclapply(exps, function(tt) calculate.gdT.score(tt, genes= genes, grep.genes = grep.genes), mc.cores = 48)
  pattern.scores = do.call(rbind, pattern.scores)
  pattern.coinfiltration = lapply(names(agg.immune.fracs.new),  function(tt)
    tryCatch( perform.lmer.coinfiltration(dt.x=agg.immune.fracs[[tt]], dt.y=pattern.scores[,frac:=patient.score], label=tt),
              error =function(e) NULL
                                  ))
  pattern.coinfiltration %<>% do.call(rbind,.) %>% 
    .[order(p)]
  list(pattern.coinfiltration = pattern.coinfiltration, pattern.scores =pattern.scores)
}

tdgv1.coinfiltration = get.coinfiltration(genes=c("TRDV1", "TRGC2"))
tdgv2.coinfiltration = get.coinfiltration(genes=c("TRDV2", "TRGC1"))


# for ( ii in seq_along(exps)) {
#   tt = exps[[ii]]
#  pattern.scores = calculate.gdT.score(tt, genes= genes, grep.genes = grep.genes)
# }
```
## Create a singleR annotation on pseudobulk. 

## Davis annotation
## Read annotations from Davis 

```{r}
library(Seurat)
files = list.files(pattern="rename_obj.rds", full.names = T, path = "/liulab/zzeng/Tigger/static", recursive = T)
metadata.davis = mclapply(files, function(file){
  sco = readRDS(file)
  label = file %>% dirname %>% basename 
  list(label =label, meta.data = sco$RNA@meta.data)
}, mc.cores = 48
)
names.metadata.davis = sapply(metadata.davis, function(tt) tt$label)
names(metadata.davis) = names.metadata.davis


davis.immune.fracs = mclapply(metadata.davis, function(tt) calcualte.immune.fraction.davis(tt$meta.data), mc.cores = 48)

agg.davis.immune.fracs = list()
for (ii in seq_along(davis.immune.fracs)) {
  immune.frac = davis.immune.fracs[[ii]]
  label.curr = metadata.davis[[ii]]$label 
  for (tt in seq_along(immune.frac)) {
    annot.curr = names(immune.frac)[[tt]]
    setnames(immune.frac[[tt]], 2, "clust")
    immune.frac[[tt]][,dataset:=label.curr][,annot:=annot.curr][,patient.new:=paste(patient, label.curr, sep = ":")]
    agg.davis.immune.fracs[[annot.curr]] %<>%  rbind(., immune.frac[[tt]])
  }
}


davis.gdT.coinfiltration = lapply(names(agg.davis.immune.fracs),  function(tt) perform.lmer.coinfiltration(dt.x=agg.davis.immune.fracs[[tt]], dt.y=gdT.scores[,frac:=patient.score], label=tt))
davis.abT.coinfiltration = lapply(names(agg.davis.immune.fracs),  function(tt) perform.lmer.coinfiltration(dt.x=agg.davis.immune.fracs[[tt]], dt.y=abT.scores[,frac:=patient.score], label=tt))
```


## Remove the cells that are positive of gdT and recalculate the fractions.


```{r}
metadata.davis.nogdt = list()
for (tt in names(metadata.davis)) {
  metadata.davis.curr = metadata.davis[[tt]]
  exps.curr = exps[[tt]]
  gdt.alleles = grep("TR[GD]",rownames(exps.curr$exp))
  try({
  inx =which( colSums(exps.curr$exp[gdt.alleles,,drop=F],na.rm = T) == 0)
  cell.inx = rownames(exps.curr$meta.data)[inx]
  metadata.davis.curr$meta.data = metadata.davis.curr$meta.data[cell.inx,]
  })
  metadata.davis.nogdt[[tt]] = metadata.davis.curr
}

davis.nogdt.immune.fracs = mclapply(metadata.davis.nogdt, function(tt) calcualte.immune.fraction.davis(tt$meta.data), mc.cores = 48)

agg.davis.nogdt.immune.fracs = list()
for (ii in seq_along(davis.nogdt.immune.fracs)) {
  immune.frac = davis.nogdt.immune.fracs[[ii]]
  label.curr = metadata.davis.nogdt[[ii]]$label 
  for (tt in seq_along(immune.frac)) {
    annot.curr = names(immune.frac)[[tt]]
    setnames(immune.frac[[tt]], 2, "clust")
    immune.frac[[tt]][,dataset:=label.curr][,annot:=annot.curr][,patient.new:=paste(patient, label.curr, sep = ":")]
    agg.davis.nogdt.immune.fracs[[annot.curr]] %<>%  rbind(., immune.frac[[tt]])
  }
}


davis.nogdt.gdT.coinfiltration = lapply(names(agg.davis.nogdt.immune.fracs),  function(tt) perform.lmer.coinfiltration(dt.x=agg.davis.nogdt.immune.fracs[[tt]], dt.y=gdT.scores[,frac:=patient.score], label=tt))
davis.nogdt.abT.coinfiltration = lapply(names(agg.davis.nogdt.immune.fracs),  function(tt) perform.lmer.coinfiltration(dt.x=agg.davis.nogdt.immune.fracs[[tt]], dt.y=abT.scores[,frac:=patient.score], label=tt))

davis.tdgv1.coinfiltration = get.coinfiltration(genes=c("TRDV1", "TRGC2"), agg.immune.fracs = agg.davis.nogdt.immune.fracs)
davis.tdgv2.coinfiltration = get.coinfiltration(genes=c("TRDV2", "TRGC1"), agg.immune.fracs = agg.davis.nogdt.immune.fracs)
davis.tdc.coinfiltration = get.coinfiltration(genes=c("TRDC"), agg.immune.fracs = agg.davis.nogdt.immune.fracs)
davis.tdv1.coinfiltration = get.coinfiltration(genes=c("TRDV1"), agg.immune.fracs = agg.davis.nogdt.immune.fracs)
davis.tdv2.coinfiltration = get.coinfiltration(genes=c("TRDV2"), agg.immune.fracs = agg.davis.nogdt.immune.fracs)
davis.TRGC1.coinfiltration = get.coinfiltration(genes=c("TRGC1"), agg.immune.fracs = agg.davis.nogdt.immune.fracs)
davis.TRGC2.coinfiltration = get.coinfiltration(genes=c("TRGC2"), agg.immune.fracs = agg.davis.nogdt.immune.fracs)


tdgv1.coinfiltration = get.coinfiltration(genes=c("TRDV1", "TRGC2"), agg.immune.fracs = agg.davis.immune.fracs)
tdgv2.coinfiltration = get.coinfiltration(genes=c("TRDV2", "TRGC1"), agg.immune.fracs = agg.davis.immune.fracs)
tdc.coinfiltration = get.coinfiltration(genes=c("TRDC"), agg.immune.fracs = agg.davis.immune.fracs)
tdv1.coinfiltration = get.coinfiltration(genes=c("TRDV1"), agg.immune.fracs = agg.davis.immune.fracs)
tdv2.coinfiltration = get.coinfiltration(genes=c("TRDV2"), agg.immune.fracs = agg.davis.immune.fracs)
TRGC1.coinfiltration = get.coinfiltration(genes=c("TRGC1"), agg.immune.fracs = agg.davis.immune.fracs)
TRGC2.coinfiltration = get.coinfiltration(genes=c("TRGC2"), agg.immune.fracs = agg.davis.immune.fracs)
out1 = grep("coinfiltration", ls(), value=T) %>% 
  sapply(., function(tt) {
    uu = eval(as.symbol(tt))
    tryCatch(uu$pattern.coinfiltration[clust=="CD8Tex"][,type:=tt], 
             error=function(e) NULL)
  }, simplify = F) %>% do.call(rbind,.)

```

## CD3+ CD8- CD4- population

```{r}
library(Seurat)
files = list.files(pattern="rename_obj.rds", full.names = T, path = "/liulab/zzeng/Tigger/static/", recursive = T)
metadata.davis = mclapply(files, function(file){
  sco = readRDS(file)
  genes.curr = rownames(sco$RNA)
  tr.alleles = grep("^TR[ABDG][VDJC]", genes.curr, value=T)
  tr.alleles = c(tr.alleles, "CD8A", "CD3E", "CD3G", "CD4", "NKG7","KLRK1") 
  label = file %>% dirname %>% basename 
  genes.sel = which(genes.curr %in% c( tr.alleles, all.genes))
  list(label =label, meta.data = sco$RNA@meta.data, exp = sco$RNA@assays$RNA@data[genes.sel,])
}, mc.cores = 48
)
names.metadata.davis = sapply(metadata.davis, function(tt) tt$label)
names(metadata.davis) = names.metadata.davis

all.annots = sapply(metadata.davis, function(tt) tt$meta.data$assign.level3_anno)%>% unlist %>% unique
all.tcell.clusters = c("TMKI67", "CD8Tem",            
"CD8Tex",
"CD4Tnaive","CD8Teff",
"Th17",
"CD8Tcm",
"Tfh",
"Th1",
"Treg",
"CD8Tnaive","CD4Tn","CD8T",
"CD4conv")

get.gdt.trgp.cd3p.cd8n.cd4n = function(metadata){
  gdt.fracs = NULL
  # browser()
  try({
    
  if(all(c("CD4", "CD8A") %in% rownames(metadata$exp))){
    tr.alleles = grep("TR[GD][VDC][0-9]*$", rownames(metadata$exp), value=T)
    trgp = metadata$exp[tr.alleles, ,drop=F] %>% colSums 
    thrs = quantile(trgp, probs=0.6)
    metadata$trgp = trgp > thrs
    cd8n.cd4n = tryCatch({
      colSums(metadata$exp[c("CD4", "CD8A"), ] == 0) ==2
      }, error=function(e) NA)
    cd3p = metadata$exp[rownames(metadata$exp) %in% c("CD3E", "CD3G"), ,drop=F] 
    if(nrow(cd3p) > 1){
      cd3p = colSums(cd3p > 0)
    }
    cd3p = cd3p > 0
    metadata$meta.data$gdt.type1 = tryCatch(cd8n.cd4n & cd3p &  metadata$trgp, 
                                            error=function(e) NULL)
    tcell.clusters = metadata$meta.data$assign.level3_anno %in% all.tcell.clusters
    metadata$meta.data$gdt.type2 = cd8n.cd4n & tcell.clusters & metadata$trgp
    metadata$meta.data$gdt.type12 = metadata$meta.data$gdt.type1 | metadata$meta.data$gdt.type2
    
    meta.data = metadata$meta.data %>% as.data.table
    col.sel = c("gdt.type1", "gdt.type2", "gdt.type12") 
    meta.data[,total.count:=.N, by=patient]
    gdt.fracs = sapply( col.sel, function(tt) meta.data[,.(frac=sum(get(tt))/unique(total.count), label=metadata$label),by=c("patient")], 
                        USE.NAMES=T, simplify = F)
  }
  })
  list(label =metadata$label, meta.data = metadata$meta.data, exp = metadata$exp, gdt.fracs=gdt.fracs)
}


metadata.davis  = sapply(metadata.davis, get.gdt.trgp.cd3p.cd8n.cd4n, simplify = F)
sapply(metadata.davis, function(tt) is.null(tt$gdt.fracs) )

metadata.davis.nogdt2 = list()
for (tt in names(metadata.davis)) {
  metadata.davis.curr = metadata.davis[[tt]]
  try({
     metadata.davis.curr$meta.data  =  metadata.davis.curr$meta.data[!metadata.davis.curr$meta.data$gdt.type12,]
  })
  metadata.davis.nogdt2[[tt]] = metadata.davis.curr
}

davis.nogdt2.immune.fracs = mclapply(metadata.davis.nogdt2, function(tt) calcualte.immune.fraction.davis(tt$meta.data), mc.cores = 48)

agg.davis.nogdt2.immune.fracs = list()
for (ii in seq_along(davis.nogdt2.immune.fracs)) {
  immune.frac = davis.nogdt2.immune.fracs[[ii]]
  label.curr = metadata.davis.nogdt2[[ii]]$label 
  for (tt in seq_along(immune.frac)) {
    annot.curr = names(immune.frac)[[tt]]
    setnames(immune.frac[[tt]], 2, "clust")
    immune.frac[[tt]][,dataset:=label.curr][,annot:=annot.curr][,patient.new:=paste(patient, label.curr, sep = ":")]
    agg.davis.nogdt2.immune.fracs[[annot.curr]] %<>%  rbind(., immune.frac[[tt]])
  }
}

davis.trgp.cd3p.cd8n.cd4n.coinfiltration = lapply(c("gdt.type1", "gdt.type2", "gdt.type12"),  function(tt){
  temp = lapply(metadata.davis, function(uu) {
    out = uu[["gdt.fracs"]][[tt]]
    if(!is.null(out)) out=out[,.(patient, frac, label)]
    out
  }
    ) 
  # browser()
  temp %<>% 
    do.call(rbind,.)
  temp[,patient.new:=paste(patient,label,sep=":")]
  
  perform.lmer.coinfiltration(dt.x=agg.davis.nogdt2.immune.fracs$assign.level3_anno, dt.y=temp, label=tt)
} )

names.curr = c("gdt.type1", "gdt.type2", "gdt.type12")
for (ii in seq_along(davis.trgp.cd3p.cd8n.cd4n.coinfiltration)) {
avi.dt = davis.trgp.cd3p.cd8n.cd4n.coinfiltration[[ii]]
p = EnhancedVolcano::EnhancedVolcano(avi.dt,
	lab = avi.dt$clust,
	x = 'beta',
	y = 'p',
	pCutoff = 1E-2,
	FCcutoff = .01,
	labSize = 4.0,
	legend=c('NS','beta','P',
		'P & beta'),
	legendPosition = 'right',
	drawConnectors = TRUE,
	widthConnectors = 0.2,
	colAlpha = 0.8,
	colConnectors = 'grey30'
	)
ggsave(sprintf("/liulab/asahu/projects/gdt/results/coinfiltration/%s.pdf", names.curr[[ii]]), p)
}

```

---

## Gene differential in specific cluster

```{r}
library(ggplot2)
library(parallel)
library(Seurat)
psuedobulk.clust.exps = mclapply(files, function(file){
  sco = readRDS(file)
  # browser()
  genes.curr = rownames(sco$RNA)
  label = file %>% dirname %>% basename 
   meta.data = sco$RNA@meta.data
  clusters.curr = meta.data$assign.level3_anno %>% unique
  psuedobulk.clust = NULL 
  try({
   nogdt.inx = !(metadata.davis[[label]]$meta.data$gdt.type12)
    psuedobulk.clust = sapply(clusters.curr, function(tt){
      inx = which((meta.data$assign.level3_anno == tt) & nogdt.inx)
      create.psuedobulk(sco$RNA@assays$RNA@data[,inx], labels = sco$RNA@meta.data$patient[inx])
    }, simplify=T) %>% 
      set_names(clusters.curr)
  }
  )
  list(label =label, meta.data = sco$RNA@meta.data, psuedobulk.clust = psuedobulk.clust, genes = genes.curr)
}, mc.cores = 60
)
names(psuedobulk.clust.exps) = sapply(psuedobulk.clust.exps, function(tt) tt$label)



davis.gdt.type12  = lapply(metadata.davis, function(uu) {
    out = uu[["gdt.fracs"]][["gdt.type12"]]
    if(!is.null(out)) out=out[,.(patient, frac, label)]
    out
  }) %>% do.call(rbind,.) %>% 
  .[,patient.new:=paste(patient, label, sep=":")]

all.genes = sapply(psuedobulk.clust.exps, function(tt) tt$genes) %>% unlist %>% unique
all.clusts = sapply(psuedobulk.clust.exps, function(tt) names(tt$psuedobulk.clust)) %>% unlist %>% unique



davis.exp.matched = sapply(all.clusts, function(clust.curr){ 
  out = matrix(NA, nrow=length(all.genes), ncol= length(davis.gdt.type12$patient.new),
               dimnames = list(all.genes, davis.gdt.type12$patient.new))
  
  for( label.curr in unique(davis.gdt.type12$label)){
    psuedobulk.curr = psuedobulk.clust.exps[[label.curr]]$psuedobulk.clust[[clust.curr]]
    if(!is.null(psuedobulk.curr)){
      patient.new.curr = paste(colnames(psuedobulk.curr), label.curr, sep=":")
      # browser()
      out[rownames(psuedobulk.curr),patient.new.curr] = psuedobulk.curr %>% as.matrix
    }
    
  }
  out
},USE.NAMES = T, simplify = F)

library(clusterProfiler)
wp2gene <- read.gmt("~/liulab_home/data/pathways/wikipathways/wikipathways-20200710-gmt-Homo_sapiens.gmt")
wp2gene <- wp2gene %>% tidyr::separate(ont, c("name","version","wpid","org"), "%")
wpid2gene <- wp2gene %>% dplyr::select(wpid, gene) #TERM2GENE
wpid2name <- wp2gene %>% dplyr::select(wpid, name) #TERM2NAME
all.genes.dt = data.table(gene=all.genes) %>% 
  .[,gene.new:=gsub("\\.[0-9]+", "", all.genes)]
gene.eg.dt = clusterProfiler::bitr(all.genes.dt$gene.new, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
all.genes.dt$ENTREZID = gene.eg.dt[match(all.genes.dt$gene.new, gene.eg.dt$SYMBOL),]$ENTREZID

deg.all = list()  
# for (clust.curr in names(davis.exp.matched)) {
require(foreach)
require(doMC)
registerDoMC(cores = 8)

# deg.all = foreach( clust.curr=names(davis.exp.matched), .inorder=T) %dopar%{
for (clust.curr in names(davis.exp.matched)) {
  
  dt.all = davis.gdt.type12[,.(x=frac,dataset=label)] 
  psuedobulk.curr = davis.exp.matched[[clust.curr]]
  avi.dt= mclapply(nrow(psuedobulk.curr) %>% seq, function(tt){
    dt.curr = dt.all%>% 
      .[,y:=psuedobulk.curr[tt,]] %>% 
      .[!is.na(y)] %>% 
      eval.nlme()
  }, mc.cores = 64) %>% 
    do.call(rbind,.) %>% data.table %>% 
    .[,gene:=all.genes.dt$gene.new]%>%
    .[,ENTREZID:=all.genes.dt$ENTREZID]
  deg.curr = avi.dt
  avi.dt = avi.dt[!grepl("^RP",gene)][!is.na(ENTREZID)][order(p)]

  curr.dir = sprintf("/liulab/asahu/projects/gdt/results/coinfiltration/%s/", clust.curr)
  dir.create(curr.dir)
  pCutoff = sort(avi.dt$p)[50]  %>% min(c(.,1E-3) ,na.rm = T)
  FCcutoff = sort(abs(avi.dt$beta),decreasing = T)[300]  %>% min(c(.,.1) ,na.rm = T)
  p = EnhancedVolcano::EnhancedVolcano(avi.dt,
                                       lab = avi.dt$gene,
                                       x = 'beta',
                                       y = 'p',
                                       pCutoff = pCutoff,
                                       FCcutoff = FCcutoff,
                                       labSize = 2.0,
                                       legend=c('NS','beta','P',
                                                'P & beta'),
                                       legendPosition = 'right',
                                       drawConnectors = TRUE,
                                       widthConnectors = 0.2,
                                       colAlpha = 0.8,
                                       colConnectors = 'grey30')
  
  ggsave(sprintf("%s/volcano.pdf", curr.dir), p)
  ## Volcano plot
  ## KEGG enrichment 
  ## GO enrichment 
  ## Msigdb enrichment
  top.genes = avi.dt[beta>0][!is.na(ENTREZID)][order(p)[1:150]]$ENTREZID
  bottom.genes = avi.dt[beta<0][!is.na(ENTREZID)][order(p)[1:150]]$ENTREZID
  kk <- clusterProfiler::enrichKEGG(gene = top.genes,
                                    organism     = 'hsa',
                                  minGSSize = 2,
                                  universe = avi.dt$ENTREZID,
                                    pvalueCutoff = 0.1)
  p = clusterProfiler::dotplot(kk, showCategory=30)
  ggsave(filename = sprintf("%s/topkegg.pdf", curr.dir),p)
  kk <- clusterProfiler::enrichGO(gene = top.genes,
                                  OrgDb    = org.Hs.eg.db,
                                  ont      = "BP",
                                  minGSSize = 2,
                                  universe = avi.dt$ENTREZID,
                                  pvalueCutoff = 0.1)
  p = clusterProfiler::dotplot(kk, showCategory=30)
  ggsave(filename = sprintf("%s/topgo_bp.pdf", curr.dir),p)
  kk <- clusterProfiler::enrichGO(gene = top.genes,
                                  OrgDb    = org.Hs.eg.db,
                                  ont      = "MF",
                                  universe = avi.dt$ENTREZID,
                                  minGSSize = 2,
                                  pvalueCutoff = 0.1)
  p = clusterProfiler::dotplot(kk, showCategory=30)
  ggsave(filename = sprintf("%s/topgo_mf.pdf", curr.dir),p)
  kk <- clusterProfiler::enrichKEGG(gene = bottom.genes,
                                    organism     = 'hsa',
                                  universe = avi.dt$ENTREZID,
                                  minGSSize = 2,
                                    pvalueCutoff = 0.1)
  p = clusterProfiler::dotplot(kk, showCategory=30)
  ggsave(filename = sprintf("%s/bottomkegg.pdf", curr.dir),p)
  kk <- clusterProfiler::enrichGO(gene = bottom.genes,
                                  OrgDb    = org.Hs.eg.db,
                                  universe = avi.dt$ENTREZID,
                                  minGSSize = 2,
                                  ont      = "BP",
                                  pvalueCutoff = 0.1)
  p = clusterProfiler::dotplot(kk, showCategory=30)
  ggsave(filename = sprintf("%s/bottomgo_bp.pdf", curr.dir),p)
  kk <- clusterProfiler::enrichGO(gene = bottom.genes,
                                  OrgDb    = org.Hs.eg.db,
                                  universe = avi.dt$ENTREZID,
                                  minGSSize = 2,
                                  ont      = "MF",
                                  pvalueCutoff = 0.1)
  p = clusterProfiler::dotplot(kk, showCategory=30)
  ggsave(filename = sprintf("%s/bottomgo_mf.pdf", curr.dir),p)
  
  avi.dt[,score:=-sign(beta)*log(p+1E-100)]
  geneList = avi.dt$score %>% 
  set_names(avi.dt$ENTREZID) %>%  
  sort(., decreasing = TRUE)
  geneList = geneList[!(names(geneList) %>% duplicated)]
  ewp <- GSEA(geneList, minGSSize =4, TERM2GENE = wpid2gene,  pvalueCutoff = 0.2, TERM2NAME = wpid2name)
  try({
    p = clusterProfiler::dotplot(ewp, showCategory=30)
  ggsave(filename = sprintf("%s/gsea_up.pdf", curr.dir),p)
  })
  
  avi.dt[,score:=sign(beta)*log(p+1E-100)]
  geneList = avi.dt$score %>% 
    set_names(avi.dt$ENTREZID) %>%  
    sort(., decreasing = TRUE)
  geneList = geneList[!(names(geneList) %>% duplicated)]
  ewp <- GSEA(geneList, minGSSize =4, TERM2GENE = wpid2gene,  pvalueCutoff = 0.2, TERM2NAME = wpid2name)
  try({
    p = clusterProfiler::dotplot(ewp, showCategory=30)
    ggsave(filename = sprintf("%s/gsea_down.pdf", curr.dir),p)
  })
  deg.all[[clust.curr]] = deg.curr
}
saveRDS(file="/liulab/asahu/projects/gdt/data/rds/coinfiltration.deg.all.genes_deg_gdtHi_vs_gdTLo.Rds", deg.all)
```



