---
title: "Analysis of gdT in immunotherapy trials"
output: html_notebook
---

# Goal : To show responders have higher gamma-delta receptors
## 
- 1. Combine cdr3 sequence from all datasets
- 2. Get meta data of response and survival 
- 3. Total number of reads in each rnaseq sample
- 4. Get cdr3 quantification of each sample (separately for TRD and TRG and  V,D,J and C)
- 5. Show responders and non-responders have significant difference in TRD and TRG 
- 6. Control for CD8B levels in the differences

```{r}
library(survminer)
library(survival)
library(magrittr)
library(tidyr)
cdr3.files = list.files("~/liulab_home/data/immunotherapy-trials/trust4/mUC_PDL1_Mariathasan", pattern = "*.tsv", full.names = T)
require(doMC)
require(foreach)
registerDoMC(cores = 48)
trust4 = foreach(cdr3.file  = cdr3.files, .inorder=T, .combine=rbind) %dopar%{
	sample.name = cdr3.file %>% basename %>% 
	gsub("GO29293_ngs_rna_targrna_", ., replacement = "") %>%
	strsplit(.,split="_") %>% unlist %>% 
	grep("^SAM", .,value=T)
	fread(cdr3.file) %>% 
	.[,sample.name:=sample.name]
}
saveRDS(file="~/liulab_home/data/immunotherapy-trials/trust4/mUC_PDL1_Mariathasan/trust4.dt.RDS", trust4)
trust4.trg = trust4[grepl("^TRG", V)|grepl("^TRG", D)| grepl("^TRG", J)|grepl("^TRG", C)] %>% 
  .[!(CDR3aa %in% c("out_of_frame", "_" ))]
trust4.trd = trust4[grepl("^TRD", V)|grepl("^TRD", D)| grepl("^TRD", J)|grepl("^TRD", C)] %>% 
  .[!(CDR3aa %in% c("out_of_frame", "_"))]
trust4.trg %<>% rbind(trust4.trd)
trust4.trg.summ  = trust4.trg[,.(total.trgs=sum(`#count`)), by=sample.name]
# trust4.trg1 = trust4.trg[!(CDR3aa %in% c("partial"))]
```

```{r}
genentech.env = local({load("/liulab/asahu/data/ssgsea/xiaoman/genentech.phenotype.RData");environment()})
surv.factors = genentech.env$aa 
surv.factors$total.trgs = trust4.trg.summ[match(rownames(surv.factors), trust4.trg.summ$sample.name),]$total.trgs
surv.factors %<>% as.data.table %>%
.[,norm.trg:=total.trgs/sizeFactor]
surv.factors[,norm.trg:=ifelse(is.na(norm.trg),0, norm.trg)]

quantize_gradient = function(tt, grades=4) {
  thrs = seq(0,1,length.out = grades+1)
 thrs = quantile(tt,probs=thrs[-c(1,length(thrs))])
  sapply(tt, function(uu) sum(thrs<uu))
}

surv.factors[,Tgd.q:=quantize_gradient(norm.trg, 4)]
surv.factors = surv.factors[Tgd.q %in% c(0,3)]
fit1 <- survfit( Surv(os, censOS) ~ Tgd.q, data = surv.factors )
p1 = ggsurvplot(fit1, surv.factors, 
                palette = "jco", pval = TRUE)
ggsave(filename="results/.figs/Mariathasan.tdg.trust4.surv.pdf", print(p1))
```