## Write input file for cdr3 prediction 
## Generate input fasta file
~/liulab_home/softwares/protein-sequence-embedding-iclr2019/data/Getz/getz.cdr3.fa

copied from : data_processing/ssgsea/imputation/simple/tgd_tcga/tcga.Rmd
and writing file copied from:
data_processing/ssgsea/imputation/scRNA/epoch-142/tcr.embedding.Rmd

```{r}
library(stringr)
seqdetpth = fread("/liulab/lsong/projects/TRUST4_TCGA/tcga_seqDepth.txt")
trust4 = fread("/liulab/lsong/projects/TRUST4_TCGA/tcga_simpleRepNoPartial.out")
process.label = function(tt){
  substring(tt,1,16) %>% gsub("-", ., replacement = ".")
}
trust4[,name:=substring(V1, 1, 16)]
filter.str = c("out_of_frame", "?")
trust4 = trust4[!(V5 %in% filter.str)] 
trust4[,myid:=paste0("V",seq(.N))]
## for PFAM header 
# >S0E8J4_GIBF5/356-387 S0E8J4.1 PF16503.4;zn-ribbon_14;
trust4[,aa.len:=stringr::str_length(V5)]
trust4[,fa.file:=paste(paste0(">",myid),  "/356-387 S0E8J4.1 PF16503.4;zn-ribbon_14;")]
trust4= trust4[!is.na(aa.len) ]
trust4.filtered = trust4[aa.len <= 25] %>% 
.[!duplicated(V5)]

trust4.filtered.trg = trust4.filtered[grepl("^TRG", V6)|grepl("^TRG", V7)| grepl("^TRG", V8)|grepl("^TRG", V9)] 
trust4.filtered.trd = trust4.filtered[grepl("^TRD", V6)|grepl("^TRD", V7)| grepl("^TRD", V8)|grepl("^TRD", V9)] 
trust4.filtered.trb = trust4.filtered[grepl("^TRB", V6)|grepl("^TRB", V7)| grepl("^TRB", V8)|grepl("^TRB", V9)] 
trust4.filtered.tra = trust4.filtered[grepl("^TRA", V6)|grepl("^TRA", V7)| grepl("^TRA", V8)|grepl("^TRA", V9)] 

trust4.trdg = rbind(trust4.filtered.trg, trust4.filtered.trd)
## map cancer types 
trust4.trdg[,bcr_patient_barcode:=substring(V1,1,12)]
clin =fread("~/liulab_home/data/tcga/TCGA_gdc/clinical_PANCAN_patient_with_followup.tsv")
trust4.trdg$cancer.type = clin[match(trust4.trdg$bcr_patient_barcode,bcr_patient_barcode)]$acronym 
trust4.trdg.all = trust4[V5 %in% unique(trust4.trdg$V5)]
trust4.trdg.all[,bcr_patient_barcode:=substring(V1,1,12)]
trust4.trdg.all$cancer.type = clin[match(trust4.trdg.all$bcr_patient_barcode,bcr_patient_barcode)]$acronym 
aa =table(trust4.trdg.all$cancer.type) 
cancer.type.sel = which(aa > 300) %>% names 
trust4.trdg.all[,cancer.type.sub:=ifelse(cancer.type %in% cancer.type.sel, cancer.type, "misc")]

trust4.trdg.all[,cdr3.type:=ifelse(grepl("^TRG", V6)|grepl("^TRG", V7)| grepl("^TRG", V8)|grepl("^TRG", V9), "TRG", "TRD")] 

trust4.trdg[,aa:=V5]
trust4.trdg.all[,aa:=V5]
```

```{r}
dir.create("/liulab/asahu/softwares/protein-sequence-embedding-iclr2019/data/tcga/")
trust4.fa = c(t(trust4.trdg[,.(fa.file, V5)]))
write.table(file = "/liulab/asahu/softwares/protein-sequence-embedding-iclr2019/data/tcga/for.gdt.cdr3.fa", x= trust4.fa, row.names = F, quote =F, col.names = F)
write.table(file = "/liulab/asahu/softwares/protein-sequence-embedding-iclr2019/data/tcga/for.gdt.cdr3.100.fa", x= trust4.fa[1:100], row.names = F, quote =F, col.names = F)
```

```{bash}
CUDA_VISIBLE_DEVICES=1  python  eval_pfam.py --models_dir /liulab/asahu/projects/icb/data/pretrained/pretrained_models/ -f  ~/liulab_home/softwares/protein-sequence-embedding-iclr2019/data/tcga/for.gdt.cdr3.fa -d -1
```

## Process the output with pfam encoder
```{r}
# source("data_processing/ssgsea/imputation/scRNA/epoch-142/source.proj.R")
source("src/from.epoch-142.source.proj.R")

color.clusters.features <- function(data, cluster,  color.cols = "condition",
                                    title="t-SNE",size=0.15, do.discrete=T, filename=NULL, normalize=TRUE, shape = 1, remove.guide=T, do.facet=F, alpha=0.5){
  require(viridis)
  require(ggthemes)
  dt1 = as.data.frame(cluster)
  colnames(dt1) = c("V1", "V2")
  ps = list()
  if(!is.null(filename)) dir.create(filename)
  for (color.col in color.cols) {
    tryCatch({
      color.col = gsub(color.col, pattern = "-", replacement = ".")
      if(normalize) data[[color.col]] = znorm(data[[color.col]])
      if(is.null(data$shape)) {
        d_cluster_1=cbind(dt1,col=data[[color.col]], shape=shape)
      }else{
        d_cluster_1=cbind(dt1,col=data[[color.col]], shape=data$shape)
      }
      title.curr = color.col
      if(title !="") title.curr = sprintf("%s_%s", title, color.col)

      do.discrete = ifelse(is.factor(d_cluster_1$col), T, F)
      p=ggplot(d_cluster_1, aes(x=V1, y=V2)) +
        geom_point(size=size,aes(color=col, shape=as.factor(shape)), alpha=alpha) +
        xlab("Dim1") + ylab("Dim2") +
        ggtitle(label = title.curr) +
        theme_classic() +
        theme(axis.text.x=element_blank(),
              axis.text.y=element_blank())

      if (do.discrete) {
        colorCount = length(unique(d_cluster_1$col))
        if(colorCount > 8){
          getPalette = colorRampPalette(brewer.pal(9, "Set1"))
          p =  p + scale_fill_manual(values = getPalette(colorCount))
        }else{
          p<- p+ scale_colour_brewer(palette = "Set2")
        }
      }else{
        p <- p+ scale_color_gradientn(colours = heat.colors(20, alpha=alpha, rev=T))
      }
      if(do.facet) p = p + facet_wrap(~col)
      if (remove.guide)  p<- p + theme(legend.position = "none")
      if(!is.null(filename)) {
        filename.curr = sprintf("%s_%s.pdf", filename, gsub(color.col, pattern="-", replacement = "_"))

        ggsave(file=filename.curr, p)
      }
      ps[[color.col]]  = p
    },
    error = function(e) ps[[color.col]]  = NA
    )

  }
  ps
}


plotUMAP <- function(data, umap.model = NULL, col = NULL, color.col = NULL,
                     title="UMAP",size=0.25, do.discrete=T, filename=NULL, n_neighbors =  15, learning_rate = 1, init = "spectral", min_dist = .01, pca = NULL,  n_threads=32, ...){
  require(uwot)
  require(ggthemes)
  require(ggplot2)
  if(is.null(umap.model)){
    set.seed(9)
    if(is.data.table(data)) data = as.data.frame(data)
    if(!is.null(col)) {
      umap.mat = as.matrix(data[,col])
    }else{
      umap.mat = as.matrix(data)
    }
    umap.model <- umap(umap.mat,
                       pca = pca, n_neighbors =  n_neighbors, learning_rate = learning_rate, init = init, min_dist = min_dist, n_threads = n_threads, ret_model=T, ...)
  }

  d_umap_1 = as.data.frame(umap.model$embedding)
  d_umap_1$col = as.factor(0)
  if(!is.null(color.col)) {
    if(length(color.col) == nrow(data)){
      d_umap_1$col=color.col
    }else{
      d_umap_1$col=data[[color.col]]
    }

    # if(is.character(color.col)){
    #   d_umap_1$col=data[[color.col]]
    # }else if(length(color.col) == nrow(data)){
    #   d_umap_1$col=color.col
    # }
  }

  do.discrete = ifelse(is.factor(d_umap_1$col), T, F)
  ## plotting the results without clustering
  p=ggplot(d_umap_1, aes(x=V1, y=V2)) +
    geom_point(size=size,aes(color=col), alpha=0.8) +
    guides(colour=guide_legend(override.aes=list(size=2))) +
    xlab("umap_1") + ylab("umap_2") +
    ggtitle(label = title) +
    theme_classic() +
    theme(axis.text.x=element_blank(),
          axis.text.y=element_blank())
  if (do.discrete) {
    p<- p+ scale_colour_brewer(palette = "Set2")
  }else{
    p <- p+ scale_color_gradient2_tableau(palette = "Orange-Blue Diverging")
    # p <- p+ scale_color_gradient2_tableau(palette = "Orange-Blue Diverging", limits=c(-.3, .3))
  }

  ##theme(legend.position = "none")
  if(!is.null(filename)) ggsave(file=filename, p)
  list(umap.model, p)
}

plot.sequence.and.clustering <- function(data.clust, text, color.col = NULL, num.plot.seq =100, text.size =2) {
  require(viridis)
  require(ggthemes)
  require(ggrepel)
  # data.tcr.clust = data.table(umap.all.p[[1]]$embedding)

  d_umap_1 = cbind(data.clust, text = text)
  if(is.null(color.col)) {
    d_umap_1$col = as.factor(0)
  }else{
    d_umap_1$col=color.col
  }
  do.discrete = ifelse(is.factor(d_umap_1$col), T, F)
  num.plot.seq = min(num.plot.seq, nrow(d_umap_1))

  p=ggplot(d_umap_1, aes(x=V1, y=V2)) +
    geom_point(aes(color=col), alpha=0.8) +
    guides(colour=guide_legend(override.aes=list(size=2))) +
    xlab("umap_1") + ylab("umap_2") +
    # ggtitle(label = title) +
    theme_classic() +
    theme(axis.text.x=element_blank(),
          axis.text.y=element_blank())

  if (do.discrete) {
    p<- p+ scale_colour_brewer(palette = "Set2")
  }else{
    p <- p+ scale_color_tableau()
  }

  p = p + geom_text_repel(
    # data.tcr = d_umap_1[sample(nrow(d_umap_1), size=100)],
    data = d_umap_1[sample(nrow(d_umap_1), size=num.plot.seq)],
    aes(x=V1, y=V2, label = text),
    size = text.size,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
  ) +
    theme(legend.position = "none")
  p
}
plot.tcr.embedding <- function(trust.curr, trust4.dt, trust4.dt.all, save.dir, title, mode="last", plot.seurat.marker=T) {
    dir.create(save.dir, recursive=T)
    
    # require(avinash)
    library(RColorBrewer)
    # trust.curr = subset.lm.embedding(trust.curr, trust4.dt, subset= trust4.dt.all$currid)
    trust.curr.curr.unique = aggregate.embedding(trust.curr, trust4.dt,  mode=mode)
    trust.curr.curr = trust.curr.curr.unique[match(trust4.dt.all$aa, trust4.dt$aa),]
    n_neighbors =15; learning_rate =1; min_dist = .01; pca = 50
    umap.all.p = plotUMAP(data = trust.curr.curr,  
                          col=NULL, color.col=as.factor(trust4.dt.all$cdr3.type), size=1, do.discrete=T, 
                          n_neighbors = n_neighbors, learning_rate = learning_rate, min_dist = min_dist, pca=pca,
                          title= title,
                          filename=NULL , n_epochs = 100, metric = "euclidean")
    
    data.umap = data.table(umap.all.p[[1]]$embedding)
    
    p1 = plot.sequence.and.clustering(data.clust=data.umap, text = trust4.dt.all$aa, color.col = as.factor(trust4.dt.all$cdr3.type), num.plot.seq =100, text.size =2) 
    ggsave(filename = sprintf("%s/cdr3_seq.pdf",save.dir), p1, width = 16, height = 16)
    xx1 = color.clusters.features( data=as.data.frame(trust4.dt.all), cluster=data.umap,  color.cols =c("cancer.type.sub", "v.allele", "d.allele", "j.allele", "c.allele"),
                                   title="type",size=2, filename= sprintf("%s/",save.dir), normalize=F, do.discrete=T, do.facet=T)

    list(umap.all.p = umap.all.p, xx1=xx1)
}

```
```{r}
filter.trs = function(xx) {
	xx[xx=="*"] = NA
	yy = table(xx) 
	aa = table(xx) 
	type.sel = which(aa > 300) %>% names 
	ifelse(xx %in% type.sel, xx, "misc") %>% 
	as.factor
}
trust4.trdg.all[,v.allele:=filter.trs(V6)] %>%
.[,d.allele:=filter.trs(V7)] %>%
.[,j.allele:=filter.trs(V8)] %>%
.[,c.allele:=filter.trs(V9)] 

trust4.trdg.all[,cancer.type.sub:=filter.trs(cancer.type)]

library(dplyr)
options(warn=-1)
modes = c("last", "mean", "middle", "SIF")
modes = c("SIF")
library(reticulate)
np <- import("numpy")
trust.files = list.files("/liulab/asahu/softwares/protein-sequence-embedding-iclr2019/data/tcga/", pattern = "for.gdt.cdr3.*.sav.fvec.npy", full.names = T)
trust.files %<>% setdiff(., grep(pattern="for.gdt.cdr3.100",.,value=T))
for (file1 in trust.files) {
    print(paste("processing", file1))
    trust4.pfam <- np$load(file1)
    
    for (mode in modes) {
        dir.name = gsub(".sav.fvec.npy", gsub("getz.cdr3.fa", basename(file1), replacement = ""), replacement = "")
        curr.dir = dirname(file1)
        
        dir.curr = sprintf("%s/outputs/%s/%s/", curr.dir, dir.name, mode)
        dir.create(dir.curr, recursive = T)
        # out = plot.tcr.embedding(trust.curr=trust4.pfam, save.dir=dir.curr, title="pfam", mode=mode)
        out1 = plot.tcr.embedding(trust.curr=trust4.pfam, trust4.dt=trust4.trdg, trust4.dt.all=trust4.trdg.all, save.dir=sprintf("%s/",dir.curr), title="pfam", mode=mode )
    }
}
options(warn=0)
```

## Determine cancer specific usuage of allele gdt receptors

```{r}

allele.agg = function(trg.mat, index=5) {
	alleles = colnames(trg.mat)
	alleles.uniq = alleles %>% substring(1,index) %>% unique
	out = lapply(alleles.uniq, function(aa){
		inx=grep(aa, alleles)
		out = trg.mat[, inx] 
		if(length(inx) > 1)  
			out %<>% rowSums(na.rm=T)
		out
	}) %>% do.call(cbind,.) %>% 
	set_colnames(alleles.uniq)

}
trust4.trgd = rbind(
 trust4[grepl("^TRG", V6)|grepl("^TRG", V7)| grepl("^TRG", V8)|grepl("^TRG", V9)], 
trust4[grepl("^TRD", V6)|grepl("^TRD", V7)| grepl("^TRD", V8)|grepl("^TRD", V9)] ) 
trust4.trgd$seqdtpth = seqdetpth[match(trust4.trgd$V1, V1)]$V2
trust4.trgd[,norm.trg:=V2*1E7/seqdtpth]

trust4.trgd[,bcr_patient_barcode:=substring(V1,1,12)]
trust4.trgd$cancertype = clin[match(trust4.trgd$bcr_patient_barcode, bcr_patient_barcode)]$acronym 



for (ii in paste0("V",6:9)) {
trust4.trgd$current.allele = substring(trust4.trgd[[ii]], 1, 5)
trust4.trgd$cancertype = as.factor(trust4.trgd$cancertype)
xx = table(trust4.trgd$current.allele) 
yy = which(xx > 200) %>% names
temp = trust4.trgd[current.allele!="*"][current.allele %in% yy]

p = ggplot(temp, aes(x=cancertype, y=norm.trg, fill=cancertype)) + 
geom_boxplot(outlier.shape = NA) +
  # geom_point(pch = 21, fill = "grey",
  #            position = position_jitter(width = 0.5, height = 0), size=0.5, alpha=0.5) +
ylim(c(0,.5)) +
   theme(axis.text.x = element_blank()) + 
  facet_wrap(~current.allele, scale="free") + 
             theme_bw() 
ggsave(filename=sprintf("results/.figs/tcga.gdt.allele.cancertype.%s.pdf", ii), p, width=16, height=10)   
}

normalize = function(mat){
	mat1 = mat[,grep("TRD", colnames(mat)), drop=F]
	if(ncol(mat1) > 1) mat1 = mat1/rowMeans(mat1)
		mat2 = mat[,grep("TRG", colnames(mat)), drop=F]
	if(ncol(mat2) > 1) mat2 = mat2/rowMeans(mat2)
		mat = cbind(mat1, mat2)
	mat[is.nan(mat)] = 0
	mat 
}

for (ii in paste0("V",6:9)) {
	trg.mat = dcast(trust4.trgd[V6!="*"], as.formula(paste("V1 ~", ii)), value.var="norm.trg", fun.aggregate=sum)
	trg.agg = trg.mat[,seq(2, ncol(trg.mat))] %>% allele.agg(index=5) %>% 
	set_rownames(trg.mat$V1)
# adding those with all zeros
	zero.samp = setdiff(seqdetpth$V1, trg.agg %>% rownames)
	trg.agg.zero = matrix(0, nrow=length(zero.samp), ncol= ncol(trg.agg)) %>%
	set_rownames(zero.samp) %>%
	rbind(trg.agg,.)
	trg.agg.melt = melt(trg.agg.zero) %>% data.table
	trg.agg.melt$Var1 %<>% substring(1,12)
	trg.agg.melt$cancertype = clin[match(trg.agg.melt$Var1, bcr_patient_barcode)]$acronym %>% 
	as.factor 
	# tr.alleles = substring(trg.agg.melt$Var2,1,5) %>% unique 
	trg.agg.melt$current.allele = substring(trg.agg.melt$Var2, 1, 5)
	xx = table(trg.agg.melt$current.allele) 
	yy = which(xx > 200) %>% names
	temp = trg.agg.melt[current.allele!="*"][current.allele %in% yy]
	
	temp1 =temp[grep("^TRD", Var2)]
	curr.tr = temp1[1]$Var2 %>% substring(1,4)
	p = ggplot(temp1, aes(x=cancertype, y=value, fill=cancertype)) + 
	geom_boxplot(outlier.shape = NA) +
	ylab("normalized TR level") +
	# ylim(c(0,.5)) +
	theme(axis.text.x = element_blank()) + 
	facet_wrap(~current.allele, scale="free") + 
	theme_bw() 
	ggsave(filename=sprintf("results/.figs/trust4/tcga.cancertype.%s.pdf", curr.tr), p, width=16, height=10)   
		try({   
	temp1 =temp[grep("^TRG", Var2)]
	curr.tr = temp1[1]$Var2 %>% substring(1,4)
	p = ggplot(temp1, aes(x=cancertype, y=value, fill=cancertype)) + 
	geom_boxplot(outlier.shape = NA) +
	ylab("normalized TR level") +
	# ylim(c(0,.5)) +
	theme(axis.text.x = element_blank()) + 
	facet_wrap(~current.allele, scale="free") + 
	theme_bw() 
	ggsave(filename=sprintf("results/.figs/trust4/tcga.cancertype.%s.pdf", curr.tr), p, width=16, height=10)
})


	trg.agg.zero.norm =  normalize(trg.agg.zero)
	trg.agg.melt = melt(trg.agg.zero.norm) %>% data.table
	trg.agg.melt$Var1 %<>% substring(1,12)
	trg.agg.melt$cancertype = clin[match(trg.agg.melt$Var1, bcr_patient_barcode)]$acronym %>% 
	as.factor 
	# tr.alleles = substring(trg.agg.melt$Var2,1,5) %>% unique 
	trg.agg.melt$current.allele = substring(trg.agg.melt$Var2, 1, 5)
	xx = table(trg.agg.melt$current.allele) 
	yy = which(xx > 200) %>% names
	temp = trg.agg.melt[current.allele!="*"][current.allele %in% yy]
	
	temp1 =temp[grep("^TRD", Var2)]
	curr.tr = temp1[1]$Var2 %>% substring(1,4)
	p = ggplot(temp1, aes(x=cancertype, y=value, fill=cancertype)) + 
	geom_boxplot(outlier.shape = NA) +
	ylab("normalized TR level") + 
	# ylim(c(0,.5)) +
	theme(axis.text.x = element_blank()) + 
	facet_wrap(~current.allele, scale="free") + 
	theme_bw() 
	ggsave(filename=sprintf("results/.figs/trust4/tcga.cancertype.%s.percent.pdf", curr.tr), p, width=16, height=10)
	try({   
	temp1 =temp[grep("^TRG", Var2)]
	curr.tr = temp1[1]$Var2 %>% substring(1,4)
	p = ggplot(temp1, aes(x=cancertype, y=value, fill=cancertype)) + 
	geom_boxplot(outlier.shape = NA) +
	ylab("normalized TR level") +
	# ylim(c(0,.5)) +
	theme(axis.text.x = element_blank()) + 
	facet_wrap(~current.allele, scale="free") + 
	theme_bw() 
	ggsave(filename=sprintf("results/.figs/trust4/tcga.cancertype.%s.percent.pdf", curr.tr), p, width=16, height=10)   
})

}

## Plot simple expression 
tcga.biolinks = readRDS("~/liulab_home/data/tcga/TCGA_gdc/tcgabiolinks/all.tcga.expression.RDS")
library(SummarizedExperiment)
tcga.mat = assay(tcga.biolinks) %>%
set_rownames(rowData(tcga.biolinks)$external_gene_name) %>% t %>% 
 scale(., scale=T, center=T) %>% t 

 grep("^TR[GD][VDJC]",rownames(tcga.mat), value=T )

```

## Expression based distribution of gamma delta allele across cancer tpes

```{r}
library(magrittr)
project.exp = readRDS("~/liulab_home/data/tcga/TCGA_gdc/tcgabiolinks/all.tcga.expression.RDS")
tcga.cancercodes = fread("~/liulab_home/data/tcga/tcga.cancercodes.txt", header=F)
library(SummarizedExperiment)
project.mat.libsize = assay(project.exp) %>% 
 colMeans
project.exp.mat = assay(project.exp) %>% sweep(., project.mat.libsize, MARGIN = 2, FUN = "/")
colnames(project.exp.mat) %<>% substring(.,1,16)
rownames(project.exp.mat) = rowData(project.exp)$external_gene_name

dt.all = colData(project.exp)
dt.curr = colData(project.exp) %>% as.data.table %>%
.[,.(days_to_death, days_to_last_follow_up, vital_status, dataset=project_id)] %>% 
.[,Event:=ifelse(vital_status=="Dead", 1, 0)]%>% 
.[,Survival:=ifelse(vital_status=="Dead", days_to_death, days_to_last_follow_up)]

trg.alleles = grep("^TR[DG]V[0-9]", rownames(project.exp.mat), value=T)
exp.curr = project.exp.mat[trg.alleles,] %>% t %>% melt %>%  
  data.table 

exp.curr$cancertype =  rep(gsub("TCGA-","", dt.all$project_id), length(trg.alleles))

avi.dt = exp.curr[!(cancertype=="THYM")]
avi.summ = avi.dt[,mean(value), by=cancertype][order(V1)]
avi.dt$cancertype %<>% factor(levels = rev(avi.summ$cancertype))
p = ggplot(avi.dt, aes(x=cancertype, y=value, fill=cancertype)) + 
geom_boxplot(outlier.shape = NA) +
   theme(axis.text.x = element_blank()) + 
  facet_wrap(~Var2) + 
ylim(c(0,.5)) +
             theme_bw() 

avi.summ$cancer = tcga.cancercodes[match(avi.summ$cancertype, V1)]$V2
avi.dt$cancer = tcga.cancercodes[match(avi.dt$cancertype, V1)]$V2%>% 
  factor(levels=avi.summ$cancer)
p1= ggplot(avi.dt, aes(x=cancer, y=value, fill=cancer)) + 
geom_boxplot(outlier.shape = NA) +
   theme(axis.text.x = element_blank()) + 
  guides(fill=FALSE)+
  coord_flip() + 
ylim(c(0,.5)) +
             theme_bw() 


```

