## Identify cell types associated 
1. SKCM samples 
2. Calculate correlation with each sample with tissue 
3. Cox regression Sur ~ g + t + gxt 
4. take significant ones only with FDR 1% 


```{r}
tcga.dataset = fread("~/liulab_home/projects/icb/data/tcga/scrna.v4.allgenes.nopcs/dataset.txt")
tcga.sample = fread("~/liulab_home/projects/icb/data/tcga/scrna.v4.allgenes.nopcs/samples_name.txt")

if.start.inx = which(colnames(tcga.dataset)== "B_cells_naive")
dataset.col = colnames(tcga.dataset)
surv.factors = tcga.dataset[,length(dataset.col) + seq(8)-8, with =F]
exp.tcga = tcga.dataset[,seq(2,if.start.inx-1), with=F] %>% as.matrix

tcga.exp = fread("~/liulab_home/data/tcga/TCGA_gdc/EBPlusPlusAdjustPANCAN_IlluminaHiSeq_RNASeqV2.geneExp.tsv")

```

```{r}
normalize.singleR.exp = function(data.name){
	ref.se =  eval(parse(text=paste0(data.name, "()"))) 
	col.dat = colData(ref.se)
	col.dat$Var1= paste(rownames(col.dat),seq(nrow(col.dat)))  
	col.dat = as.data.table(col.dat)
	norm.ref.se = assay(ref.se) %>%
	t() %>% scale(., center=T, scale=T) %>% t()
	rownames(norm.ref.se) %<>% toupper
	colnames(norm.ref.se) = col.dat$Var1
	list(ref.se=ref.se, norm.ref.se=norm.ref.se)
}

library(SingleR)
singleR.data.names = grep("Data$", ls("package:SingleR"), value=T)
singleR.dataset = lapply(singleR.data.names, function(ii) normalize.singleR.exp(data.name=ii))
names(singleR.dataset) = singleR.data.names
saveRDS(file="~/liulab_home/data/single_cell/singleR.dataset.RDS", singleR.dataset)

```

```{r}

skcm.inx = which(tcga.dataset$cancertype =="SKCM")
surv.factors.skcm = surv.factors[skcm.inx]
sample.names.skcm = tcga.sample$x[skcm.inx]
exp.tcga.skcm = exp.tcga[skcm.inx,] %>%
  scale(.,scale=T, center=T) %>% 
  set_rownames(tcga.sample$x[skcm.inx]) %>% 
  t
run.seq = seq(2, if.start.inx-1)
genes = colnames(tcga.dataset)[run.seq]

calculate.singleR.corelation <- function(mat, ref.data) {
	col.dat = colData(ref.data$ref.se)
	col.dat = as.data.table(col.dat)
	colnames(ref.data$norm.ref.se) = col.dat$label.fine
	## defined in avinash.scrna
	calcCorEnrich(ref.data$norm.ref.se, mat)
}

singleR.dataset = readRDS("~/liulab_home/data/single_cell/singleR.dataset.RDS")

cor.out = lapply(singleR.dataset, function(ii) calculate.singleR.corelation(mat=exp.tcga.skcm, ref.data=ii)) 

skcm.cor.out = do.call(rbind,cor.out)
cell.types =  rownames(skcm.cor.out) %>% unique
skcm.cor.unique = sapply( cell.types, function(tt){
	inx = which(rownames(skcm.cor.out) == tt)
	if(length(inx) > 1){
		out = skcm.cor.out[inx,] %>% colMeans(na.rm=T)
	}else{
		out = skcm.cor.out[inx,]
	}
	out
} ) %>% t

skcm.cor.gdt = skcm.cor.unique["T_cell:gamma-delta", ,drop=F]  
skcm.cor.gdt %<>% rbind(skcm.cor.gdt)
gdt.cor.out = lapply(singleR.dataset, function(ii) calculate.singleR.corelation(mat=skcm.cor.gdt, ref.data=ii))

```

## Identify variable genes to reduce calculations 
```{r}
library(Seurat)
curr.ref = singleR.dataset[["HumanPrimaryCellAtlasData"]]
exp.gamma.delta =  curr.ref$norm.ref.se[,curr.ref$ref.se@colData$label.fine=="T_cell:gamma-delta"] 

tt = curr.ref$ref.se %>% assay %>%
FindVariableFeatures

tt <- tt[order(tt$vst.variance.standardized, decreasing = TRUE), , drop = FALSE]
variable.features = rownames(tt) %>% head(2000)
genes.sel = intersect(variable.features, rownames(exp.tcga.skcm))
exp.tcga.skcm = exp.tcga.skcm[genes.sel,]
exp.gamma.delta = exp.gamma.delta[toupper(genes.sel),]
```

```{r}
library(survival)
calculate.cox.interaction = function(dtx){
	tryCatch(
	{
		dtx = dtx[!is.na(gene)]
		aa =  coxph(Surv(Survival, Event) ~ gene + cell.type + gene*cell.type, dtx)
		summary(aa)$coefficients[,c(1,5)] %>% t%>% c
	},
	error=  function(e) rep(NA,6)
	)

}

library(parallel)
interaction.out = mclapply(cell.types, function(cc){ 
	dt.curr = surv.factors.skcm[,.(Survival=OS.time,Event=OS.filtered)]%>%
	.[,cell.type:=skcm.cor.unique[cc,]]
	out = apply(exp.tcga.skcm,1, function(exp.curr) {
		dt.curr[,gene:=exp.curr]%>%
		calculate.cox.interaction
	})
	t(out)
}, mc.cores = 60) 

interaction.mat = interaction.out %>% sapply(., function(tt) tt[,5]) %>% t

enrichment.p = WGCNA::corAndPvalue(t(interaction.mat), exp.gamma.delta, 
             use = "pairwise.complete.obs", 
             alternative = "two.sided",
             method="spearman")

avi.dt = data.table(cell.type = cell.types, P=enrichment.p$p[,1],
	correlation = enrichment.p$cor[,1])
avi.dt[,p.adj:=p.adjust(P, method="fdr")]
library(EnhancedVolcano)
p = EnhancedVolcano(avi.dt,
	lab = avi.dt$cell.type,
	x = 'correlation',
	y = 'p.adj',
	pCutoff = 1e-40,
	FCcutoff = .1,
	pointSize = c(ifelse(avi.dt$p.adj< 1E-40, 2, 1)),
	labSize = 4.0,
	legend=c('NS','Correlation','Adj.P value',
		'Adj.P value & correlation'),
	legendPosition = 'bottom',
	legendLabSize = 8,
	legendIconSize = 4.0,
	drawConnectors = TRUE,
	widthConnectors = 0.2,
	xlim = c(-.5, .5),
	colAlpha = 0.8,
	colConnectors = 'grey30'
	)
ggsave(file="results/.figs/gdt.dyfn.pdf", p, width=16, height=10)
		
```

## Check high TRD high Fibroblast have high IL-17 and low IFNG
```{r}

dt.curr = colData(project.exp) %>% as.data.table %>%
.[,.(days_to_death, days_to_last_follow_up, vital_status, sample, dataset=project_id)] %>% 
.[,Event:=ifelse(vital_status=="Dead", 1, 0)]%>% 
.[,Survival:=ifelse(vital_status=="Dead", days_to_death, days_to_last_follow_up)]

surv.factors.skcm = dt.curr[dataset == "TCGA-SKCM",] %>% 
  .[match(colnames(skcm.cor.out), sample)]
exp.tcga.skcm = project.skcm[,colnames(skcm.cor.out)]

surv.factors.skcm[,TRD.exp:=exp.tcga.skcm["TRGV9",]] %>% 
.[,Fibroblasts:=skcm.cor.unique["Fibroblasts activated",]]
surv.factors.skcm[,TRD.level:=ifelse(TRD.exp > mean(TRD.exp), "TRD-high", "TRD-low")] %>%
.[,Fibroblast.level:=ifelse(Fibroblasts > mean(Fibroblasts), "Fibroblast-high", "Fibroblast-low")] 
grp.1 = which(
  (surv.factors.skcm$TRD.level =="TRD-high") & 
  (surv.factors.skcm$Fibroblast.level =="Fibroblast-high")) 
  
grp.2 = which(
  (surv.factors.skcm$TRD.level =="TRD-low") & 
  (surv.factors.skcm$Fibroblast.level =="Fibroblast-high")) 

wilcox.test(exp.tcga.skcm["IFNG",grp.1], exp.tcga.skcm["IFNG", grp.2], alternative = "greater")
grep("IFNG", rownames(exp.tcga.skcm), value=T)
ils = grep("IL17", rownames(exp.tcga.skcm), value=T)
for (il in ils) {
  print(il)
print(wilcox.test(exp.tcga.skcm[il,grp.1], exp.tcga.skcm[il, grp.2], alternative = "less"))
}

```




## Download data from TCGAbiolinks 

```{r}
setwd("~/liulab_home/data/tcga/TCGA_gdc/tcgabiolinks/")
library(TCGAbiolinks)

query <- GDCquery(project = "TCGA-SKCM",
                  data.category = "Transcriptome Profiling",
                  data.type = "Gene Expression Quantification", 
                  workflow.type = "HTSeq - FPKM-UQ")
GDCdownload(query)
# data <- GDCprepare(query)
skcm.exp <- GDCprepare(query, 
                      save = TRUE, 
                      summarizedExperiment = TRUE, 
                      save.filename = "SKCMHTSeq_FPKM-UQ.rda")
skcm.exp.mat = assay(skcm.exp) %>%
set_rownames(rowData(skcm.exp)$external_gene_name) %>% t %>% 
 scale(., scale=T, center=T) %>% t 
colnames(skcm.exp.mat) %<>% substring(.,1,16)
barcode.skcm = gsub("\\.", sample.names.skcm, replacement="-")
setdiff(barcode.skcm, colnames(skcm.exp.mat))
skcm.exp.mat = skcm.exp.mat[,barcode.skcm]
```

## Check with TRG and TRD alleles 
```{r}
load("~/liulab_home/data/tcga/TCGA_gdc/tcgabiolinks/SKCMHTSeq_FPKM-UQ.rda")
skcm.exp = data; rm(data)

exp.tcga.skcm1 = exp.tcga[skcm.inx,] %>%
  scale(.,scale=T, center=T) %>% 
  set_rownames(tcga.sample$x[skcm.inx]) %>% 
  t() 
exp.tcga.skcm.trg = exp.tcga.skcm1[c("TRDV3", "TRGV5"),] 
exp.tcga.skcm.trg  %<>% rbind(., TRDG=rowSums(.))

interaction.trg.out = mclapply(cell.types, function(cc){ 
	dt.curr = surv.factors.skcm[,.(Survival=OS.time,Event=OS.filtered)]%>%
	.[,cell.type:=skcm.cor.unique[cc,]]
	out = apply(exp.tcga.skcm.trg,1, function(exp.curr) {
		dt.curr[,gene:=exp.curr]%>%
		calculate.cox.interaction
	})
	t(out)
}, mc.cores = 60) 

interaction.trg.p = interaction.trg.out %>% sapply(., function(tt) tt[,6]) %>% t
interaction.trg.cor = interaction.trg.out %>% sapply(., function(tt) tt[,5]) %>% t


library(survminer)
surv.factors.skcm[,TRD.exp:=exp.tcga.skcm.trg["TRDV3",]] %>% 
.[,Fibroblasts:=skcm.cor.unique["Fibroblasts activated",]]
surv.factors.skcm[,TRD.level:=ifelse(TRD.exp > mean(TRD.exp), "TRD-high", "TRD-low")] %>%
.[,Fibroblast.level:=ifelse(Fibroblasts > mean(Fibroblasts), "Fibroblast-high", "Fibroblast-low")] 

fit1 <- survfit( Surv(OS.time, OS.filtered) ~ TRD.level, data = surv.factors.skcm )
p = ggsurvplot_facet( fit1, surv.factors.skcm, facet.by = "Fibroblast.level", 
                palette = "jco", pval = TRUE)
ggsave(filename="results/.figs/gdt.fibo.dyfn.pdf", p)


surv.factors.skcm[,TRD.exp:=exp.tcga.skcm.trg["TRDV3",]] %>% 
.[,Myelocytes:=skcm.cor.unique["Myelocyte",]]
surv.factors.skcm[,TRD.level:=ifelse(TRD.exp > mean(TRD.exp), "TRD-high", "TRD-low")] %>%
.[,Myelocyte.level:=ifelse(Myelocytes > mean(Myelocytes), "Myelocyte-high", "Myelocyte-low")] 

fit1 <- survfit( Surv(OS.time, OS.filtered) ~ TRD.level, data = surv.factors.skcm )
p = ggsurvplot_facet( fit1, surv.factors.skcm, facet.by = "Myelocyte.level", 
                palette = "jco", pval = TRUE)
ggsave(filename="results/.figs/gdt.Myelocyte.dyfn.pdf", p)


surv.factors.skcm[,TRD.exp:=exp.tcga.skcm.trg["TRDV3",]] %>% 
.[,B_Cell:=skcm.cor.unique["Exhausted B cells",]] %>%
.[,gdT:=skcm.cor.unique["T_cell:gamma-delta",]] 

surv.factors.skcm[,TRD.level:=ifelse(TRD.exp > mean(TRD.exp), "TRD-high", "TRD-low")] %>%
.[,B_Cells.level:=ifelse(B_Cell > mean(B_Cell), "B_Cells-high", "B_Cells-low")] 

fit1 <- survfit( Surv(OS.time, OS.filtered) ~ B_Cells.level, data = surv.factors.skcm )
p = ggsurvplot_facet( fit1, surv.factors.skcm, facet.by = "TRD.level", 
                palette = "jco", pval = TRUE)
ggsave(filename="results/.figs/gdt.B_Cells.dyfn.pdf", p)


surv.factors.skcm[,Memory_B_Cells:=skcm.cor.unique["Memory B-cells",]]
surv.factors.skcm[,gdT.level:=ifelse(gdT > mean(gdT), "gdT-high", "gdT-low")] %>%
.[,Memory_B_Cells.level:=ifelse(Memory_B_Cells > mean(Memory_B_Cells), "Memory_B_Cells-high", "Memory_B_Cells-low")] 

fit1 <- survfit( Surv(OS.time, OS.filtered) ~ Memory_B_Cells.level, data = surv.factors.skcm )
p = ggsurvplot_facet( fit1, surv.factors.skcm, facet.by = "gdT.level", 
                palette = "jco", pval = TRUE)
ggsave(filename="results/.figs/gdt.Memory_B_Cells.gdt.dyfn.pdf", p)
```

## Analysis of TIDE

```{r}
mycorAndP = function(x,y, use = "pairwise.complete.obs", 
             alternative = "two.sided",
             method="spearman"){
	rownames(x) %<>% toupper
	rownames(y) %<>% toupper
	common = intersect(rownames(x), rownames(y))
	x = x[common, , drop=F]
	y = y[common, , drop=F]
 WGCNA::corAndPvalue(x, y, 
             use = use, 
             alternative = alternative, 
             method=method)
}
mycor = function(x,y, use = "pairwise.complete.obs", 
             method="spearman"){
	rownames(x) %<>% toupper
	rownames(y) %<>% toupper
	common = intersect(rownames(x), rownames(y))
	x = x[common, , drop=F]
	y = y[common, , drop=F]
 WGCNA::cor(x, y, 
             use = use, 
             method=method)
}
tide.skcm = fread("~/liulab_home/data/TIDE/Results/Dysfunction_scores/raw_output/TCGA.SKCM.RNASeq.OS_base.interaction")
gene.df <- clusterProfiler::bitr(tide.skcm$V1, toType="SYMBOL", fromType="ENTREZID", OrgDb="org.Hs.eg.db")
tide.skcm$gene=gene.df[match(tide.skcm$V1,gene.df$ENTREZID),]$SYMBOL
tide.skcm[,logp:=-sign(z)*log(p)]
tide.score = tide.skcm$z %>% as.matrix(ncol=1) %>% 
set_rownames(tide.skcm$gene)
mycor.out = lapply(singleR.dataset, function(tt) 
	mycorAndP(tt$norm.ref.se, tide.score)
)
mycor.out = lapply(names(mycor.out), function(tt){
	lapply(mycor.out[[tt]], c) %>% do.call(cbind,.) %>% as.data.table  %>% 
	.[,cell.type:=singleR.dataset[[tt]]$ref.se@colData$label.fine]%>% 
	.[,p.adj:=p.adjust(p, method="fdr")]
}) %>% do.call(rbind, .)

tide.sub.skcm = tide.skcm[order(p)[1:200]]
tide.sub.score = tide.sub.skcm$logp %>% as.matrix(ncol=1) %>% 
set_rownames(tide.sub.skcm$gene)
mycor.out.sub = lapply(singleR.dataset, function(tt) 
	mycorAndP(tt$norm.ref.se, tide.sub.score, method="spearman")
)
mycor.out.sub1 = lapply(names(mycor.out.sub[c(2,4,5)]), function(tt){
	lapply(mycor.out.sub[[tt]], c) %>% do.call(cbind,.) %>% as.data.table  %>% 
	.[,cell.type:=singleR.dataset[[tt]]$ref.se@colData$label.fine]
}) %>% do.call(rbind, .) %>% 
	.[,p.adj:=p.adjust(p, method="fdr")] %>% 
.[order(p.adj)] %>% 
.[,list(p=min(p), cor=mean(cor), p.adj=min(p.adj)), by=cell.type]
library(EnhancedVolcano)
mycor.out.sub2 =mycor.out.sub1[!grep("regulatory", cell.type)]
p = EnhancedVolcano(mycor.out.sub2,
	lab = mycor.out.sub2$cell.type,
	x = 'cor',
	y = 'p.adj',
	pCutoff = 5e-2,
	FCcutoff = .1,
	pointSize = c(ifelse(mycor.out.sub2$p.adj< 5e-2, 2, 1)),
	labSize = 4.0,
	legend=c('NS','Correlation','Adj.P value',
		'Adj.P value & correlation'),
	legendPosition = 'bottom',
	legendLabSize = 8,
	legendIconSize = 4.0,
	drawConnectors = TRUE,
	widthConnectors = 0.2,
	xlim = c(-.3, .3),
	ylim = c(0, 3),
	colAlpha = 0.8,
	colConnectors = 'grey30'
	)
ggsave(file="results/.figs/gdt.t_cell.dysfunction.pdf", p, width=16, height=10)


calculate.singleR.corelation <- function(mat, ref.data) {
	col.dat = colData(ref.data$ref.se)
	col.dat = as.data.table(col.dat)
	colnames(ref.data$norm.ref.se) = col.dat$label.fine
	mycor(ref.data$norm.ref.se, mat, method="pearson")
}

skcm.exp.singer.cor = lapply(singleR.dataset[c(2,4,5)], function(ii) calculate.singleR.corelation(mat=skcm.exp.mat, ref.data=ii)) 
skcm.exp.singer.cor = skcm.cor.unique

surv.curr = surv.factors.skcm 
cyto.genes = c("CD8A", "CD8B", "GZMA", "GZMB", "PRF1")
surv.curr[,cytotoxic:=colMeans(skcm.exp.mat[cyto.genes,])]
surv.curr[,"Monocytes_CD14":=skcm.exp.singer.cor ["Monocytes, CD14+",]]
surv.curr[,cytotoxic.level:=ifelse(cytotoxic > median(cytotoxic), "cytotoxic-high", "cytotoxic-low")] %>%
.[,Monocytes_CD14.level:=ifelse(Monocytes_CD14 > median(Monocytes_CD14), "Monocytes_CD14-high", "Monocytes_CD14-low")] 

fit1 <- survfit( Surv(OS.time, OS.filtered) ~ cytotoxic.level, data = surv.curr )
p = ggsurvplot_facet( fit1, surv.curr, facet.by = "Monocytes_CD14.level", 
                palette = "jco", pval = TRUE)
ggsave(filename="results/.figs/cytoxic.dyfn.Monocytes_CD14.pdf", p)

surv.curr[,"Tgd.vg3":=skcm.exp.singer.cor ["Tgd (Tgd.VG3+24AHI)",]]
surv.curr[,cytotoxic.level:=ifelse(cytotoxic > median(cytotoxic), "cytotoxic-high", "cytotoxic-low")] %>%
.[,Tgd.vg3.level:=ifelse(Tgd.vg3 > median(Tgd.vg3), "Tgd.vg3-high", "Tgd.vg3-low")] 

fit1 <- survfit( Surv(OS.time, OS.filtered) ~ cytotoxic.level, data = surv.curr )
p = ggsurvplot_facet( fit1, surv.curr, facet.by = "Tgd.vg3.level", 
                palette = "jco", pval = TRUE)
ggsave(filename="results/.figs/cytoxic.dyfn.Tgd.vg3.pdf", p)


surv.curr[,"Macrohage.MF":=skcm.exp.singer.cor ["Macrophages (MF.11CLOSER)",]]
surv.curr[,cytotoxic.level:=ifelse(cytotoxic > median(cytotoxic), "cytotoxic-high", "cytotoxic-low")] %>%
.[,Macrohage.MF.level:=ifelse(Macrohage.MF > median(Macrohage.MF), "Macrohage.MF-high", "Macrohage.MF-low")] 

fit1 <- survfit( Surv(OS.time, OS.filtered) ~ cytotoxic.level, data = surv.curr )
p = ggsurvplot_facet( fit1, surv.curr, facet.by = "Macrohage.MF.level", 
                palette = "jco", pval = TRUE)
ggsave(filename="results/.figs/cytoxic.dyfn.Macrohage.MF.pdf", p)



surv.curr[,"Stem_Cells":=skcm.exp.singer.cor ["Stem cells (SC.CMP.DR)",]]
surv.curr[,cytotoxic.level:=ifelse(cytotoxic > median(cytotoxic), "cytotoxic-high", "cytotoxic-low")] %>%
.[,Stem_Cells.level:=ifelse(Stem_Cells > median(Stem_Cells), "Stem_Cells-high", "Stem_Cells-low")] 

fit1 <- survfit( Surv(OS.time, OS.filtered) ~ cytotoxic.level, data = surv.curr )
p = ggsurvplot_facet( fit1, surv.curr, facet.by = "Stem_Cells.level", 
                palette = "jco", pval = TRUE)
ggsave(filename="results/.figs/cytoxic.dyfn.Stem_Cells.pdf", p)

```


## Visualize correlation of correlation matrix 

```{r}

M<-cor(t(skcm.cor.unique[1:40,]), method="spearman")
head(round(M,2))
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
library(corrplot)
pdf("results/.figs/cor.cor.pdf", width=16, height=10)
corrplot(M, method="color", col=col(200),  
         type="upper", order="hclust", 
         addCoef.col = NULL,
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         # p.mat = p.mat, sig.level = 0.01, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )
dev.off()

## 21 and 1
which(M>0.85 & M < .88,arr.ind=T) 
sel = rownames(M)[c(6,11)]
aa = singleR.dataset[[1]]
bb = colData(aa$ref.se) %>% as.data.table
bb$inx = seq(nrow(bb))
bb.sel = bb[label.fine %in% sel] ## 1 and 3d
avi.dt = aa$norm.ref.se[,c(14,20)] %>% as.data.table %>% 
set_colnames(c("cd4T", "Bcells"))
p = ggplot(avi.dt, aes(x=cd4T, y=Bcells)) + geom_point(size=.5, alpha=0.5) + 
theme_bw()
ggsave(filename="results/.figs/Neutrophils.Megakaryocytes.pdf", p) 
```


## Evaluate the gerber correlation 

```{r}
gerber.correlation = function(hist.returns, lookback = nrow(hist.returns), threshold = 0.5, psuedo.count=0) {
    n <- ncol(hist.returns)
    nperiods <- nrow(hist.returns)
  
    if (lookback > nperiods) lookback <- nperiods
    index <- (nperiods - lookback + 1) : nperiods
  
    standard.deviation <- apply(hist.returns[index,,drop=F], 2, sd, na.rm = T)
    threshold <- threshold * standard.deviation
  
    correlation <- matrix(1, n, n, dimnames=list(colnames(hist.returns), colnames(hist.returns)))
    for (i in 1:(n-1))
        for (j in 2:n) {
            pos <- sum((hist.returns[,i] >= threshold[i] & hist.returns[,j] >= threshold[j]) |
                       (hist.returns[,i] <= -threshold[i] & hist.returns[,j] <= -threshold[j]), na.rm = T) + psuedo.count
      
            neg <- sum((hist.returns[,i] >= threshold[i] & hist.returns[,j] <= -threshold[j]) |
                       (hist.returns[,i] <= -threshold[i] & hist.returns[,j] >= threshold[j]), na.rm = T) + psuedo.count
      
            correlation[i,j] <- correlation[j,i] <- (pos - neg) / (pos + neg)
        }
    correlation  
}

gerber.correlation2 = function(x, y, threshold.pos = 0.5, threshold.neg=0.5, psuedo.count=0) {
    nx <- ncol(x)
    ny <- ncol(y)
    nperiods <- nrow(x)

  
  
    standard.deviation.x <- apply(x, 2, sd, na.rm = T)
    standard.deviation.y <- apply(y, 2, sd, na.rm = T)
    thresholdx.pos <- threshold.pos * standard.deviation.x
    thresholdy.pos <- threshold.pos * standard.deviation.y
    thresholdx.neg <- threshold.neg * standard.deviation.x
    thresholdy.neg <- threshold.neg * standard.deviation.y
  
    correlation <- matrix(1, nx, ny, dimnames=list(colnames(x), colnames(y)))
    for (i in 1:nx)
        for (j in 1:ny) {
            pos <- sum((x[,i] >= thresholdx.pos[i] & y[,j] >= thresholdy.pos[j]) |
                       (x[,i] <= -thresholdx.pos[i] & y[,j] <= -thresholdy.pos[j]), na.rm = T) + psuedo.count
      
            neg <- sum((x[,i] >= thresholdx.neg[i] & y[,j] <= -thresholdy.neg[j]) |
                       (x[,i] <= -thresholdx.neg[i] & y[,j] >= thresholdy.neg[j]), na.rm = T) + psuedo.count
      
            correlation[i,j] <- (pos - neg) / (pos + neg)
        }
    correlation  
}

curr.label = singleR.dataset[[1]]$ref.se@colData$label.fine 
sel = sapply(unique(curr.label), function(tt){
	inx = which(curr.label %in% tt)
	if(length(inx) >=2){
		inx = inx[1:2]
	}else{
		inx=NULL
	}
}) 
sel = sel[1:20] %>% unlist

temp.exp = singleR.dataset[[1]]$norm.ref.se[,sel] %>% 
set_colnames(curr.label[sel])
out.gerber = gerber.correlation(temp.exp) %>% 
set_rownames(colnames(temp.exp)) %>%
set_colnames(colnames(temp.exp))
out.spearman = cor(temp.exp, method="spearman", use = "pairwise.complete.obs")
out.pearson = cor(temp.exp, method="pearson", use = "pairwise.complete.obs")
# ord = hclust(out.spearman)
hr <- hclust(as.dist(1-out.spearman), method="complete") 
out.gerber=out.gerber[hr$order, hr$order] 
out.spearman=out.spearman[hr$order, hr$order] 
out.pearson=out.pearson[hr$order, hr$order] 

pdf("results/.figs/gerber.comparison.pdf", width=10, height=7)
par(mfrow=c(1,2))
corrplot(out.gerber, title = "gerber", 
	method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(out.gerber),
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )
aa = corrplot(out.spearman, title = "spearman", 
	method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(out.spearman),
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )
# corrplot(out.pearson, title = "pearson", 
# 	method="color", col=col(200),  
#          type="upper", order="original", 
#          addCoef.col = T,
#           # Add coefficient of correlation
#          tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
#          diag=FALSE 
#          )

dev.off()

seqs = seq(0.2, 3, 0.25) 

pdf("results/.figs/gerber.comparison.pdf", width=10, height=7)
par(mfrow=c(3,4))
for (ii in seqs) {
temp = gerber.correlation(temp.exp, threshold = ii) %>% 
.[hr$order, hr$order]
corrplot(temp, title = as.character(ii), 
	method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(temp.exp),
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )
}
dev.off()

#' Calculate Preferential Expression Measure
#' Please refer Kryuchkova-Mostacci, N., & Robinson-Rechavi, M. (2017). A benchmark of gene expression tissue-specificity metrics. Briefings in Bioinformatics, 18(2), 205–214. https://doi.org/10.1093/bib/bbw008
#' Defined as  log10(expr(g,t) * sum_{g,t} expr/(sum_{t}expr * sum_{g}expr ))
#' @param expr expression value
#'
#' @export
#' @return
calculate.pem = function(expr){
	sum.expr = sum(expr,na.rm=T)
	cols = matrix(colSums(expr,na.rm=T), nrow=1)
	rows = matrix(rowSums(expr,na.rm=T), ncol=1)
	aa = log10((expr+1) * sum.expr/( rows %*% cols))
	aa[is.nan(aa)] = NA
	aa[is.infinite(aa)] = NA
	aa
}

calculate.tau = function(expr){
	expr = expr/apply(expr,1,max,na.rm=T)
	rowSums(1-expr,na.rm=T)/(ncol(expr)-1) 
}

calculate.tau.unique = function(expr, labels){
	expr.unique = lapply(unique(labels), function(tt){
		inx =which(labels==tt)
		out = expr[,inx]
		if(length(inx) > 1) out = rowMeans(out,na.rm=T)
	}) %>% do.call(cbind,.)
 calculate.tau( expr.unique)

}
expr =  singleR.dataset[[1]]$ref.se %>% assay
tau = calculate.tau.unique(expr, singleR.dataset[[1]]$ref.se@colData$label.fine)


pdf("results/.figs/tissue.specfic.hist.pdf")
hist(tau, 200)
dev.off()
temp.exp.ts = temp.exp *  tau

out.gerber = gerber.correlation2(temp.exp.ts, temp.exp)
out.spearman = cor(temp.exp.ts, temp.exp, method="spearman", use = "pairwise.complete.obs")
out.pearson = cor(temp.exp.ts, temp.exp, method="pearson", use = "pairwise.complete.obs")
hr <- hclust(as.dist(1-out.spearman), method="complete") 

pdf("results/.figs/tissue.specfic.gerber.comparison.pdf", width=10, height=7)
par(mfrow=c(1,3))
corrplot(out.gerber, title = "gerber", 
	method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(out.gerber),
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )
aa = corrplot(out.spearman, title = "spearman", 
	method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(out.spearman),
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )
corrplot(out.pearson, title = "pearson", 
	method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(out.pearson),
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )

dev.off()


pem = calculate.pem(expr)
pdf("results/.figs/tissue.specfic.hist.pdf")
hist(pem[,2], 200)
dev.off()

pem.sub = pem[,sel] %>% set_colnames(colnames(temp.exp))
out.gerber = gerber.correlation2(pem.sub, temp.exp)
out.spearman = cor(pem.sub, temp.exp, method="spearman", use = "pairwise.complete.obs")
out.pearson = cor(pem.sub, temp.exp, method="pearson", use = "pairwise.complete.obs")

pdf("results/.figs/tissue.specfic.gerber.comparison.pem.pdf", width=10, height=7)
par(mfrow=c(1,3))
corrplot(out.gerber, title = "gerber", 
	method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(out.gerber),
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )
aa = corrplot(out.spearman, title = "spearman", 
	method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(out.spearman),
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )
corrplot(out.pearson, title = "pearson", 
	method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(out.pearson),
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )
dev.off()

seqs = seq(0.2, 3, 0.25) 

pdf("results/.figs/gerber.comparison.pdf", width=10, height=7)
par(mfrow=c(3,4))
for (ii in seqs) {
temp = gerber.correlation2(pem.sub, temp.exp, threshold = ii, psuedo.count=1)
corrplot(temp, title = as.character(ii), 
	method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(temp.exp),
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )
}
dev.off()

pdf("results/.figs/gerber.comparison.pem.asymmetric.threshold.pdf", width=10, height=7)
temp = gerber.correlation2(pem.sub, temp.exp, threshold.pos =0.5, threshold.neg=0.5, psuedo.count=0)
corrplot(temp, title = "0.5;05", 
	method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(temp.exp),
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )

temp = gerber.correlation2(pem.sub, temp.exp, threshold.pos =1.5, threshold.neg=0.75, psuedo.count=0)
corrplot(temp, title = "2;0.5", 
	method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(temp.exp),
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )
dev.off()


calc.gerber = function(x,y, threshold.pos =.75, threshold.neg=0.75) {
	require(magrittr)
	rownames(x) %<>% toupper
	rownames(y) %<>% toupper
	common = intersect(rownames(x), rownames(y))
    gerber.correlation2(x[common,],y[common,],threshold.pos =threshold.pos, threshold.neg=threshold.neg)
}




tcga.ref.gerber = calc.gerber(exp.tcga.skcm, pem.sub, threshold.pos =1, threshold.neg=1)
pdf("results/.figs/gerber.tcga.pdf") 
hist(tcga.ref.cor,breaks=200)
dev.off()
tcga.ref.gerber %<>% t

skcm.cor.out.sub = skcm.cor.out[sel,] %>% set_rownames(colnames(pem.sub))

tcga.cor = cor(tcga.ref.gerber %>% t)
skcm.cor.sub = cor(skcm.cor.out.sub %>% t)

pdf("results/.figs/gerber.comparison.skcm.pdf")
corrplot(tcga.cor, title = "gerber", 
	method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(tcga.cor),
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )
corrplot(skcm.cor.sub, title = "correlation", 
	method="color", col=col(200),  
         type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(temp.exp),
          # Add coefficient of correlation
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )
dev.off()

```
## tau vs. z-score 
## Gerber tau 
## Gerber z-score
## Spearman tau
## Spearman z-score
```{r}

calc.cor = function(x,y, method="spearman") {
	require(magrittr)
	rownames(x) %<>% toupper
	rownames(y) %<>% toupper
	common = intersect(rownames(x), rownames(y))
    cor(x[common,],y[common,], method=method, use = "pairwise.complete.obs") 
}

skcm.gerber.tau = calc.gerber(exp.tcga.skcm, pem.sub, threshold.pos =1, threshold.neg=1)
skcm.gerber.z = calc.gerber(exp.tcga.skcm, temp.exp, threshold.pos =1, threshold.neg=1)
skcm.spearman.tau = calc.spearman(exp.tcga.skcm, pem.sub)
skcm.spearman.z = calc.cor(exp.tcga.skcm, temp.exp)

temp.250.exp = temp.exp %>% apply(.,2, function(tt){
	xx = abs(tt)
	thrs = sort(xx,decreasing=T)[1000]
	ifelse(xx> thrs, tt,NA)
})
skcm.pearson.200.z = calc.cor(exp.tcga.skcm, temp.250.exp, method="spearman")

var.names  = c( "skcm.gerber.z", "skcm.spearman.tau", "skcm.spearman.z", "skcm.pearson.200.z")
make.corrplot = function(var.name, ...){
	var = eval(parse(text = var.name)) %>% 
	cor(., method="spearman")
	corrplot( var, 
		title = var.name, 
		method="color", col=col(200),  
		type="upper", order="original", 
         addCoef.col = T,
         number.cex= 7/ncol(var),
         tl.cex=0.5, tl.col="black", tl.srt=45, #Text label color and rotation
         diag=FALSE 
         )
}


pdf("results/.figs/speaman.comparison.spearman.skcm.pdf"); 
par( mfrow = c(2, 2))
for (var.name in var.names) {
	make.corrplot(var.name)
}; dev.off()

```


## Identify tissue by tissue interactions 

```{r}

exp.tcga.norm = exp.tcga %>%
  scale(.,scale=T, center=T) %>% 
  set_rownames(tcga.sample$x) %>% 
  t
run.seq = seq(2, if.start.inx-1)
genes = colnames(tcga.dataset)[run.seq]

calculate.singleR.corelation <- function(mat, ref.data) {
	col.dat = colData(ref.data$ref.se)
	col.dat = as.data.table(col.dat)
	colnames(ref.data$norm.ref.se) = col.dat$label.fine
	## defined in avinash.scrna
	calcCorEnrich(ref.data$norm.ref.se, mat)
}
cor.out = lapply(singleR.dataset, function(ii) calculate.singleR.corelation(mat=exp.tcga.norm, ref.data=ii)) 

tcga.cor.out = do.call(rbind,cor.out)
cell.types =  rownames(tcga.cor.out) %>% unique
tcga.cor.unique = sapply( cell.types, function(tt){
	inx = which(rownames(tcga.cor.out) == tt)
	if(length(inx) > 1){
		out = tcga.cor.out[inx,] %>% colMeans(na.rm=T)
	}else{
		out = tcga.cor.out[inx,]
	}
	out
} ) %>% t
saveRDS(file="~/liulab_home/data/tcga/tcga.singler.Rds", tcga.cor.unique)
```

```{r}
tcga.cor.gdt = tcga.cor.unique["T_cell:gamma-delta",]  
avi.dt = data.table(gamma.delta.score=tcga.cor.gdt, cancer.type=tcga.dataset$cancertype)
avi.dt.summ= avi.dt[,.(mean(gamma.delta.score, na.rm=T)), by=cancer.type][order(V1)]
avi.dt[,cancer.type:=factor(cancer.type, levels=avi.dt.summ$cancer.type)]
p = ggplot(avi.dt, aes(x=cancer.type, y=gamma.delta.score)) + 
geom_boxplot() +
  geom_point(pch = 21, fill = "grey",
             position = position_jitter(width = 0.5, height = 0)) +
             theme_bw()
ggsave(filename="~/tcga.gdt.scores.pdf", p, width=16, height=10)        
```

## Interaction with gdt alleles 
 <!-- grep("^TRDV", rownames(project.exp.mat), value=T) -->
<!-- [1] "TRDV1" "TRDV2" "TRDV3" -->

```{r}
library(coxme)
library(data.table)
library(magrittr)
library(parallel)
library(survival)

extract_coxme_table <- function (mod){
    beta <- mod$coefficients #$fixed is not needed
    nvar <- length(beta)
    nfrail <- nrow(mod$var) - nvar
    se <- sqrt(diag(mod$var)[nfrail + 1:nvar])
    z<- round(beta/se, 2)
    p<- signif(1 - pchisq((beta/se)^2, 1), 2)
    table=data.frame(cbind(beta,se,z,p))
    return(table)
}

eval.coxme = function(dat, dtx){
    tryCatch(
        {
            dtx$col = dat
            dtx = dtx[!is.na(col)]
            dtx = dtx[dataset %in% names(which(table(dtx$dataset) > 10))] 
            aa =  coxme(Surv(Survival, Event) ~ col + (1|dataset), dtx)
            extract_coxme_table(aa)
        },error=  function(e) rep(NA,4))
}

calculate.cox.interaction.pancancer.tr = function(dtx){
    tryCatch(
    {
        dtx = dtx[!is.na(gene)]
        dtx = dtx[dataset %in% names(which(table(dtx$dataset) > 10))]
        dtx.summ =dtx[,.(var(gene)), by=dataset][V1 >1E-10]
        dtx = dtx[dataset %in% dtx.summ$dataset]
        aa =  coxme(Surv(Survival, Event) ~ gene + trg + gene*trg + (1|dataset), dtx)
        tt = extract_coxme_table(aa)
        tt["gene:trg",]
    },
    error=  function(e) rep(NA,4),
    warning = function(e) rep(NA,4)
    )
}


library(magrittr)
project.exp = readRDS("~/liulab_home/data/tcga/TCGA_gdc/tcgabiolinks/all.tcga.expression.RDS")
library(SummarizedExperiment)
project.exp.mat = assay(project.exp) %>%
set_rownames(rowData(project.exp)$external_gene_name) %>% t %>% 
 scale(., scale=T, center=T) %>% t 
colnames(project.exp.mat) %<>% substring(.,1,16)

dt.curr = colData(project.exp) %>% as.data.table %>%
.[,.(days_to_death, days_to_last_follow_up, vital_status, dataset=project_id)] %>% 
.[,Event:=ifelse(vital_status=="Dead", 1, 0)]%>% 
.[,Survival:=ifelse(vital_status=="Dead", days_to_death, days_to_last_follow_up)]

dt.skcm = dt.curr[dataset == "TCGA-SKCM",] 
project.skcm  = project.exp.mat[,dt.curr$dataset=="TCGA-SKCM"]

trg.alleles = grep("^TRDV", rownames(project.exp.mat), value=T)

trg.interaction.out = lapply(trg.alleles, function(trg.allele){ 
    dt.curr[,trg:=project.exp.mat[trg.allele,]]
    dataset.sel = sapply(unique(dt.curr$dataset), function(tt){
    	out = F
    	try({
    	out = coxph(Surv(Survival, Event) ~ trg, data=dt.curr[dataset %in% tt]) %>% summary 
    	xx = out$coefficient
    	ifelse(xx[1] < 0 & xx[5] < 0.05, T, F)
    })
    }) %>% which %>% names
    inx.sel = which(dt.curr$dataset %in% dataset.sel)
    exp.sel = project.exp.mat[,inx.sel]
    dt.sel = dt.curr[inx.sel,]
    out = mclapply(seq(nrow(exp.sel)), function(ii) {
        dt.sel[,gene:=exp.sel[ii,]]%>%
        calculate.cox.interaction.pancancer.tr
    }, mc.cores = 60) %>% 
    do.call(rbind, .) %>%
    as.data.table %>% 
    set_colnames(c("estimate","se", "z", "P")) %>% 
    .[,gene:=rownames(exp.sel)]
}) 

xx = project.exp.mat[grep("^TR[GD][VDJC]", rownames(project.exp.mat), value=T),]
yy = project.exp.mat[c("BTN3A1" ,"BTN2A1"),]
aa = cor(t(xx), t(yy))

```
```{r}
plot.outliers.scatter =  function(avi.dt) {
  require(ggrepel)
  aa = fit.loess(avi.dt$V1, avi.dt$V2)
  inx.pos = (avi.dt$V2 - aa$fitted) %>% order(decreasing=T) %>% 
    .[1:40]
  inx.neg = (aa$fitted- avi.dt$V2 ) %>% order(decreasing=T) %>% 
    .[1:30]
  # avi.dt.sub = avi.dt[inx]
  avi.dt.sub = avi.dt[unique(c(
    order(avi.dt$V1,decreasing=F)[1:30],
    order(avi.dt$V2,decreasing=F)[1:30],
    order(avi.dt$V1,decreasing=T)[1:40],
    order(avi.dt$V2,decreasing=T)[1:40], 
    inx.pos,
    inx.neg
  ))]
  p.sc.volcano = ggplot(data=avi.dt, aes(x=V1, y= V2))  + 
    geom_point(alpha=0.5) +
    theme_minimal(base_size = 12) + theme(legend.position = "bottom") +
    geom_text_repel(
      data = avi.dt.sub,
      aes(x = V1, y = V2, label = gene),    
      size = 3,
      box.padding = unit(0.35, "lines"),
      point.padding = unit(0.3, "lines")
    ) 
  p.sc.volcano
}
# find.outliers <- function(x, y, limits = c(-Inf, Inf), ...) {
#   lo <- loess.smooth(x, y
#   fx <- approxfun(lo$x, lo$y * limits[1L])
#   fy <- approxfun(lo$x, lo$y * limits[2L])
# 
#   which(y < fx(x) | y > fy(x))
# }
fit.loess <- function(x, y, loess.spn=0.3, ...) {
	dt = data.table(x=x,y=y)
  fit <- loess(
      formula = y~ x,
      data = dt,
      span = loess.spn
    )
  fit
}


```

## Distribution of gamma delta alleles 
<!-- -1. Sum of TRD  -->
<!-- -2. Sum of TRG -->
<!-- -3. Sum of TRDV etc.  -->
<!-- -4. Individual genes  -->

```{r}
library(tidyverse)
logNormalize <- function(counts, scale.factor=10000,
                                   use.log= TRUE) {
  colsum = colSums(counts) 
  out = t(counts)/colsum * scale.factor 
  if(use.log) out %<>% log1p
  t(out)
}
plot.geombox = function(exp.mat, cancertypes, variable, facet=NULL){
  # if(nrow(exp.mat) >1) facet = 
  avi.dt = exp.mat %>% t %>% 
    set_rownames(cancertypes) %>% melt %>% data.table()
  setnames(avi.dt, 1, "cancer.type")
  avi.dt.summ= avi.dt[,.(mean(value, na.rm=T)), by=cancer.type][order(V1)]
  avi.dt[,cancer.type:=factor(cancer.type, levels=avi.dt.summ$cancer.type)]
	p = ggplot(avi.dt, aes(x=cancer.type, y=value)) + 
	geom_boxplot(outlier.shape = NA) +
	theme_bw()
	if(ncol(exp.mat) >1) {
		p = p + facet_wrap(~Var2,scales = "free")
	}
	p = p + theme(axis.text.x = element_text(angle = 90))
	ggsave(filename=sprintf("/liulab/asahu/projects/gdt/results/distribution/%s.pdf",variable), p, width=16, height=10) 
	p
}

project.exp.tpm = logNormalize(assay(project.exp))
project.exp.tpm %<>% set_rownames(rowData(project.exp)$external_gene_name)

vars = c("TRDV", "TRGV", "TRD[VDJC]", "TRG[VDJC]")
vars = c("TRAV")
for (Var in vars) {
  trg.alleles = grep(Var, rownames(project.exp.mat), value=T)
exp.mat = project.exp.tpm[trg.alleles[1:10],]
  p = plot.geombox(exp.mat, cancertype=dt.curr$dataset, variable = Var)
}

	 # p + facet_wrap(~Var2) + theme(axis.text.x = element_text(angle = 90))
```


## Interaction in melanoma only
```{r}

	library(survival)
lsos = function (..., n = 10)
{
    .ls.objects(..., order.by = "Size", decreasing = TRUE, head = TRUE,
        n = n)
}

calculate.cox.interaction.tr = function(dtx){
    tryCatch(
    {
        dtx = dtx[!is.na(gene)]
        aa =  coxph(Surv(Survival, Event) ~ gene + trg +  gene*trg , dtx)%>%
      summary 
    	xx = aa$coefficient
        xx["gene:trg",]
    },
    error=  function(e) rep(NA,5),
    warning = function(e) rep(NA,5)
    )
}

# inx.skcm = which(dt.curr$dataset == "TCGA-SKCM")

# dt.skcm=dt.curr[inx.skcm,] %>% scale(.,center=T, scale=T)
# trg.alleles = grep("^TRGV", rownames(project.exp.mat), value=T)

# project.exp ## SKCM specific
# project.exp.mat = assay(project.exp) %>%
# set_rownames(rowData(project.exp)$external_gene_name) %>% t %>% 
#  scale(., scale=T, center=T) %>% t 
# colnames(project.exp.mat) %<>% substring(.,1,16)

# dt.skcm = colData(project.exp) %>% as.data.table %>%
# .[,.(days_to_death, days_to_last_follow_up, vital_status, dataset=project_id)] %>% 
# .[,Event:=ifelse(vital_status=="Dead", 1, 0)]%>% 
# .[,Survival:=ifelse(vital_status=="Dead", days_to_death, days_to_last_follow_up)]

trg.alleles = grep("^TR[GD][VDJC]", rownames(project.exp.mat), value=T) 
xx=setdiff(  grep("^TRGV", rownames(project.exp.mat),value=T), "TRGV9")
project.skcm= rbind(project.skcm, 
                      "TRDV2GV9" =colSums(project.skcm[c("TRDV2","TRGV9"),]),
                      "TRGV9minus" =colSums(project.skcm[xx,]),
                      "TRBV" =colSums(project.skcm[grep("^TRBV", rownames(project.exp.mat),value=T),])
                     )
trg.alleles = c("TRDV1", "TRDV2", "TRDV2GV9", "TRGV9", "TRGV9minus", "TRBV") 

# c( c("TRDV1", "TRDV2", "TRGV1", "TRGV9", "TRGV6", "TRGV10", "TRGC1", "TRGC2") 

trg.interaction.out.skcm = lapply(trg.alleles, function(trg.allele){ 
    dt.skcm[,trg:=project.skcm[trg.allele,]]%>%
      .[,CD8B:=project.skcm["CD8B",]]
    out = mclapply(seq(nrow(project.skcm)), function(ii) {
        dt.skcm[,gene:=project.skcm[ii,]]%>%
        calculate.cox.interaction.tr
    }, mc.cores = 60) %>% 
    do.call(rbind, .) %>%
    data.table %>% 
    set_colnames(c("estimate", "exp", "se", "z", "P")) %>% 
    .[,gene:=rownames(project.skcm)]
    out 
}) 

names(trg.interaction.out.skcm) = trg.alleles
# trg.interaction.skcm.dt = lapply(trg.interaction.out.skcm, function(tt){
# tt%<>% data.table 
# tt$gene=rownames(project.skcm)
# tt
# })


trg.cox.out = lapply(trg.alleles, function(trg.allele){ 
      dt.skcm[,trg:=project.skcm[trg.allele,]]%>%
      .[,CD8B:=project.skcm["CD8B",]]
      xx = rep(NA,5)
      try({
      	out = coxph(Surv(Survival, Event) ~ trg, data=dt.skcm) %>% summary 
      	xx = out$coefficient["trg",]

      })
      xx
    }) %>% do.call(rbind, .) %>%
    data.table %>% 
    set_colnames(c("estimate", "exp", "se", "z", "P")) %>% 
    .[,gene:=trg.alleles]


trg.int.dt = lapply(names(trg.interaction.out.skcm), function(uu)
{
	tt = trg.interaction.out.skcm[[uu]]
	
	tt[,fdr:=p.adjust(P,method="fdr")]
	tt[,up_or_down:=ifelse(z > 0 , "enhance", "dysfunction")] %>% 
	.[,label:=uu]
	tt = tt[fdr<0.1]
	if(nrow(tt) > 200) tt = tt[order(abs(z), decreasing=T)[1:200]]
	tt
}) %>% do.call(rbind,.)

gene.eg.dt = clusterProfiler::bitr(rownames(project.skcm), fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db") %>%
data.table(.) %>%
setkey(., SYMBOL) 

trg.int.dt[,ENTREZID:= gene.eg.dt[match(trg.int.dt$gene, SYMBOL)]$ENTREZID]
trg.int.dt = trg.int.dt[!is.na(ENTREZID)][label%in% c()]


library(clusterProfiler)
aa = trg.interaction.out.skcm$TRGC1[!is.na(z)]
aa[,ENTREZID:= gene.eg.dt[match(aa$gene, SYMBOL)]$ENTREZID]
aa = aa[!is.na(ENTREZID)][order(z)]

geneList = (-1 *aa$z) %>% 
set_names(aa$ENTREZID) %>%  
sort(., decreasing = TRUE)
geneList = geneList[!(names(geneList) %>% duplicated)]
# ego3 <- gseGO(geneList     = geneList,
#               OrgDb        = org.Hs.eg.db,
#               ont          = "CC",
#               nPerm        = 1000,
#               minGSSize    = 100,
#               maxGSSize    = 500,
#               pvalueCutoff = 0.05,
#               verbose      = FALSE)

ego3 <- gseKEGG(geneList     = geneList,
               organism     = 'hsa',
               minGSSize    = 120,
               pvalueCutoff = 0.05,
               verbose      = FALSE)

 p = dotplot(ego3, showCategory=30)
 ggsave("results/.figs/skcm.trg.pathway.enhance.TRGC1.pdf", p)

library(clusterProfiler)
geneLists = sapply(trg.interaction.out.skcm, function(aa){
	bb = aa[!is.na(z),]
	bb$ENTREZID= gene.eg.dt[match(bb$gene, SYMBOL)]$ENTREZID
	bb = bb[!is.na(ENTREZID)][order(z)]
	geneList = bb$z %>% 
	set_names(bb$ENTREZID) %>%  
	sort(., decreasing = TRUE) 
	# geneList = bb[order(z,decreasing=T)]$ENTREZID
	geneList
})
geneLists = geneLists[sapply(geneLists, length) > 0]
geneLists = geneLists[sort(names(geneLists))]
xx <- myCompareCluster(geneLists, fun="gseKEGG", 
			 organism     = 'hsa',
               minGSSize    = 120,
               pvalueCutoff = 0.01,
               verbose      = FALSE)

p <-  dotplot.compareClusterResult(xx, by="enrichmentScore", font.size=8, showCategory=7) + theme(axis.text.x = element_text(angle = 45))
ggsave(file="results/.figs/all.trgs.skcm.pathway.pdf", p, width=12, height=10)

## Overlap with murad hits 
trdv2vg9.interaction = trg.interaction.out.skcm$TRDV2GV9
library(readxl)
library(tidyr)
library(readxl)
crispr = data.table(read_xlsx("~/liulab_home/projects/gdt/data/Murad_Screen_Data.xlsx", sheet = 1))
gene.eg.dt = clusterProfiler::bitr(unique(crispr$id), fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
crispr$ENTREZID = gene.eg.dt[match(crispr$id, gene.eg.dt$SYMBOL),]$ENTREZID
crispr.overlap = crispr %>% 
  .[,gene:=toupper(id)] %>% 
  .[gene%in% toupper(crispr$id)]
crispr.overlap$interaction.score = trdv2vg9.interaction[match(crispr.overlap$gene, gene)]$z
# genes.overlap = intersect(crispr$id %>% toupper, trdv2vg9.interaction$gene %>% toupper)
p =   ggplot(crispr.overlap, aes(x=interaction.score, y=`neg|lfc`)) + 
  geom_point(alpha=0.5)

crispr.overlap.sel = crispr.overlap[`neg|lfc` < -3 & interaction.score < -1.25]

 kk <- clusterProfiler::enrichKEGG(gene = crispr.overlap.sel$ENTREZID,
                                   universe = crispr.overlap$ENTREZID,
                                   minGSSize = 4,
                                    organism     = 'hsa',
                                    pvalueCutoff = 0.05)
 library(clusterProfiler)
 
 
 kk <- clusterProfiler::enrichGO(gene = crispr.overlap.sel$ENTREZID, universe = crispr.overlap$ENTREZID,minGSSize = 4,OrgDb= org.Hs.eg.db, 
                                 ont= "BP", pvalueCutoff = 0.05)
dotplot(kk, showCategory=50)

## Comparison between TRDV1 and TRDV2 
TRBV.interaction = trg.interaction.out.skcm$TRBV
TRDV2.interaction = trg.interaction.out.skcm$TRDV2
TRDV2.interaction$TRBV.interactions = TRBV.interaction[match(TRDV2.interaction$gene, gene)]$z
TRBV.TRDV2.interaction = TRDV2.interaction[,.(V1=TRBV.interactions, V2=z, gene=gene)]
p = plot.outliers.scatter(TRBV.TRDV2.interaction)
p1 = p + xlab("TRBV") + ylab("TRDV2") + theme_bw()
gene.eg.dt = clusterProfiler::bitr(unique( TRBV.TRDV2.interaction$gene), fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
TRBV.TRDV2.interaction$ENTREZID = gene.eg.dt[match(TRBV.TRDV2.interaction$gene,gene.eg.dt$SYMBOL),]$ENTREZID
TRBV.TRDV2.interaction = TRBV.TRDV2.interaction[!is.na(ENTREZID)]
trdv.interaction.sel = TRBV.TRDV2.interaction[V1< -1.8 & V2>1.8]
kk <- clusterProfiler::enrichGO(gene = trdv.interaction.sel$ENTREZID, universe = TRBV.TRDV2.interaction$ENTREZID,minGSSize = 4,OrgDb= org.Hs.eg.db, ont= "BP", pvalueCutoff = 0.05)

dotplot(kk, showCategory=50)

## plot specific examples of Kerantizing genes 

out = dt.skcm 
avi.dt = out %>%
.[,TRGV9:=project.skcm["TRGV9",]] %>% 
.[,TRGV9:=ifelse(TRGV9> median(TRGV9)," high", " low")] %>% 
.[,ELF5:=as.numeric(project.skcm["ELF5",])] %>% 
.[,ELF5:=ifelse(ELF5> median(ELF5),"ELF5 high", "ELF5 low")]

fit1 <- survfit( Surv(Survival, Event) ~ ELF5, data = avi.dt)
p = ggsurvplot_facet( fit1, avi.dt, facet.by = "TRGV9",  palette = "jco", pval = TRUE)

fit1 <- survfit( Surv(Survival, Event) ~TRGV9 , data = avi.dt)
p1 = ggsurvplot_facet( fit1, avi.dt, facet.by = "ELF5",  palette = "jco", pval = TRUE)
```



## Pathway enrichment
```{r}
library(plyr)

dotplot.compareClusterResult <- function(object, x=~Cluster, colorBy="p.adjust",
                                         showCategory=5, by="geneRatio",
                                         split=NULL, includeAll=TRUE, font.size=12, title="") {

    # df <- fortify(object, showCategory=showCategory, by=by,
                  # includeAll=includeAll, split=split)
	df = as.data.frame(object)
    my.plotting.clusterProfile(df, x=x, type="dot", colorBy=colorBy,
                            by=by, title=title, font.size=font.size)
}

my.plotting.clusterProfile <- function(clProf.reshape.df,
                                    x = ~Cluster,
                                    type = "dot",
                                    colorBy = "p.adjust",
                                    by = "geneRatio",
                                    title="",
                                    font.size=12) {
    Description <- Percentage <- Count <- Cluster <- GeneRatio <- p.adjust <- pvalue <- NULL # to satisfy codetools
    require(DOSE)
    if (type == "bar") {
        if (by == "percentage") {
            p <- ggplot(clProf.reshape.df,
                        aes(x=Description, y = Percentage, fill=Cluster))
        } else if (by == "count") {
            p <- ggplot(clProf.reshape.df,
                        aes(x=Description, y = Count, fill=Cluster))
        } else {

        }
        p <- p +
            geom_bar() +
                coord_flip()
    }
    if (type == "dot") {
        if (by == "rowPercentage") {
            p <- ggplot(clProf.reshape.df,
                        aes_(x = x, y = ~Description, size = ~Percentage))
        } else if (by == "count") {
            p <- ggplot(clProf.reshape.df,
                        aes_(x = x, y = ~Description, size = ~Count))
        } else if (by == "geneRatio") {
            p <- ggplot(clProf.reshape.df,
                        aes_(x = x, y = ~Description, size = ~GeneRatio))
        } else {
        clProf.reshape.df$Cluster = factor(clProf.reshape.df$Cluster, levels=sort(as.character(unique(clProf.reshape.df$Cluster))))
        # browser()	
            p <- ggplot(clProf.reshape.df,
                        aes_(x = x, y = ~Description, size = 0.3))
        }
        if (any(colnames(clProf.reshape.df) == colorBy)) {
            p <- p +
                geom_point() +
                aes_string(color=colorBy) +
                scale_color_continuous(low="red", high="blue", guide=guide_colorbar(reverse=TRUE))
            ## scale_color_gradientn(guide=guide_colorbar(reverse=TRUE), colors = sig_palette)
        } else {
            p <- p + geom_point(colour="steelblue")
        }
    }
    p <- p + xlab("") + ylab("") + ggtitle(title) +
        theme_dose(font.size)
    ## theme(axis.text.x = element_text(colour="black", size=font.size, vjust = 1)) +
    ##     theme(axis.text.y = element_text(colour="black",
    ##           size=font.size, hjust = 1)) +
    ##               ggtitle(title)+theme_bw()
    ## p <- p + theme(axis.text.x = element_text(angle=angle.axis.x,
    ##                    hjust=hjust.axis.x,
    ##                    vjust=vjust.axis.x))
    return(p)
}

extract_params <- function(x) {
    y <- rlang::quo_text(x)
    if (is.function(x)) y <- sub('\nNULL$', '', y)

    y <- gsub('"', '', y) %>%
        sub(".*\\(", "", .) %>%
        sub("\\)", "", .) %>%
        gsub("\\s+", "", .)

    y <- strsplit(y, ",")[[1]]
    params <- sub("=.*", "", y)
    vals <- sub(".*=", "", y)
    i <- params != vals
    params <- params[i]
    vals <- vals[i]
    names(vals) <- params
    return(as.list(vals))
}

myCompareCluster = function (geneClusters, fun = "enrichGO", data = "", ...){
	is.gse = substring(fun, 1,3) == "gse"
    if (is.character(fun)) {
        fun <- eval(parse(text = fun))
    }
    if (typeof(geneClusters) == "language") {
        if (!is.data.frame(data)) {
            stop("no data provided with formula for compareCluster")
        }
        else {
            genes.var = all.vars(geneClusters)[1]
            grouping.formula = gsub("^.*~", "~", as.character(as.expression(geneClusters)))
            geneClusters = dlply(.data = data, formula(grouping.formula),
                .fun = function(x) {
                  as.character(x[[genes.var]])
                })
        }
    }
    # if(!is.gse){
    	clProf <- llply(geneClusters, .fun = function(i) {
    		x = NULL 
    		try({
    			x = suppressMessages(fun(i, ...))
    		if (class(x) == "enrichResult" || class(x) == "groupGOResult" || class(x) == "gseaResult") {
    			x = as.data.frame(x)
    		}
    	})
    		x
    	})
    # }else{
    # 	clProf <- lapply(geneClusters, function(i) {
    # 		x = suppressMessages(fun(i, ...))
    # 		as.data.frame(x)
    # 	})

    # }
    clusters.levels = names(geneClusters)
    clProf.df <- ldply(clProf, rbind)
    # browser()
    if(is.gse)  clProf.df$geneRatio = clProf.df$enrichmentScore
    if (nrow(clProf.df) == 0) {
        stop("No enrichment found in any of gene cluster, please check your input...")
    }
    clProf.df <- plyr::rename(clProf.df, c(.id = "Cluster"))
    clProf.df$Cluster = factor(clProf.df$Cluster, levels = clusters.levels)
    if (is.data.frame(data) && grepl("+", grouping.formula)) {
        groupVarName <- strsplit(grouping.formula, split = "\\+") %>%
            unlist %>% gsub("~", "", .) %>% gsub("^\\s*", "",
            .) %>% gsub("\\s*$", "", .)
        groupVars <- sapply(as.character(clProf.df$Cluster),
            strsplit, split = "\\.") %>% do.call(rbind, .)
        for (i in seq_along(groupVarName)) {
            clProf.df[, groupVarName[i]] <- groupVars[, i]
        }
        i <- which(colnames(clProf.df) %in% groupVarName)
        j <- (1:ncol(clProf.df))[-c(1, i)]
        clProf.df <- clProf.df[, c(1, i, j)]
    }
    res <- new("compareClusterResult", compareClusterResult = clProf.df,
        geneClusters = geneClusters, .call = match.call(expand.dots = TRUE))
    params <- modifyList(extract_params(args(fun)), extract_params(res@.call))
    keytype <- params[["keyType"]]
    if (is.null(keytype))
        keytype <- "UNKNOWN"
    readable <- params[["readable"]]
    res@keytype <- keytype
    res@readable <- as.logical(readable)
    res@fun <- params[["fun"]]
    return(res)
}
	```

## Read GO pathways 
```{r}
library(survminer)
library(ggplot2)
 go.pathways = qusage::read.gmt("~/liulab_home/data/pathways/download.baderlab.org/EM_Genesets/current_release/Human/symbol/Human_GOBP_AllPathways_with_GO_iea_June_01_2020_symbol.gmt")
 search.inx1 = c("Cholesterol", 'vitamin k', "Coenzyme Q", "steroid", "Mevalonate", "LDL", "fatty acid")
 search.inx2 = c("lipid", "endothelial", "pattern", "heat shock", "natural killer", "toll-like")
 search.inx = c(search.inx1, search.inx2)
pathway.sel = search.inx %>% sapply(., function(tt) 
	grep(pattern=tt, names(go.pathways), value=T, ignore.case=T)
	)
pathway.curr = pathway.sel[search.inx1] %>% unlist
genes.curr = go.pathways[pathway.curr] %>% unlist %>% unique
go.pathways$"CHOLESTEROL BIOSYNTHESIS PATHWAY%WIKIPATHWAYS_20200510%WP197%HOMO SAPIENS"

aa = trg.interaction.out.skcm$TRGV9[order(abs(z), decreasing=T)]

nn =  grep("BTN3A1", go.pathways)
genes.curr =  go.pathways$"T CELL RECEPTOR SIGNALING PATHWAY%GOBP%GO:0050852" %>% sort
aa[gene %in% genes.curr]


out = dt.skcm 
avi.dt = out %>%
.[,TRGC1:=project.skcm["TRGC1",]] %>% 
.[,TRGC1:=ifelse(TRGC1> median(TRGC1),"high", "low")] %>% 
.[,IGLV3:=as.numeric(project.skcm["IGLV3-13",])] %>% 
.[,IGLV3:=ifelse(IGLV3> median(IGLV3),"high", "low")]

fit1 <- survfit( Surv(Survival, Event) ~ IGLV3, data = avi.dt)
p = ggsurvplot_facet( fit1, avi.dt, facet.by = "TRGC1",  palette = "jco", pval = TRUE)
ggsave("results/.figs/skcm.TRGC1.IGLV3-13.pdf")

fit1 <- survfit( Surv(Survival, Event) ~ TRGC1, data = avi.dt)
p = ggsurvplot_facet( fit1, avi.dt, facet.by = "IGLV3",  palette = "jco", pval = TRUE)
ggsave("results/.figs/skcm..IGLV3-13.TRGC1.pdf")

avi.dt = trg.interaction.out.skcm$TRGC1
tcga.surv.genes = readRDS("~/liulab_home/data/tcga/tcga.cox.genes.R")
avi.dt = avi.dt[gene %in% tcga.surv.genes$OS$genes]
library(EnhancedVolcano)
p = EnhancedVolcano(avi.dt,
	lab = avi.dt$gene,
	x = 'estimate',
	y = 'P',
	pCutoff = 4E-5,
	FCcutoff = .1,
	pointSize = c(ifelse(avi.dt$fdr <  .03, 2, 1)),
	labSize = 4.0,
	legend=c('NS','Estimate','P value',
		'P value & Estimate'),
	legendPosition = 'bottom',
	legendLabSize = 8,
	legendIconSize = 4.0,
	drawConnectors = TRUE,
	widthConnectors = 0.2,
	xlim = c(-.5, .5),
	colAlpha = 0.8,
	colConnectors = 'grey30'
	)
ggsave(file="results/.figs/trgc1.gene.interaction.pdf", p, width=10, height=8)

```

```{r}

get.enrich = function(genes, universe,
	organism     = 'hsa',
		pvalueCutoff = 0.05, ...) {
	if( !sum(grepl("^[0-9]", universe)) > 100 ){
		universe = clusterProfiler::bitr(universe, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
		genes =  universe[match(genes,universe$SYMBOL),]$ENTREZID
		universe = universe$ENTREZID
	}

	kk <- clusterProfiler::enrichKEGG(gene         = genes, 
		universe = universe,
		organism     = organism,
		pvalueCutoff = pvalueCutoff,
		...
		)
	clusterProfiler::dotplot(kk, showCategory=30) 
}
```

## Correlate between markers of TLS/GCR and gdT
```{r}
library(ggplot2)
library(ggrepel)
## Identify  TLS genes correled with TRD allele levels. 
tls.genes = c("MS4A1", "CD40", "CXCR5", "CXCL13", "CCR7", "CCL19",  "TFRC", "CD8B", "PDCD1",  "CTLA4","HAVCR2", "CD274", "PDCD1LG2")
setdiff(tls.genes, rownames(project.skcm))
grep("TRBV", rownames(project.skcm), value=T)
temp = c("TRDV2", "TRGV9", "TRBV2")
trg.alleles = grep("^TR[GD]V",rownames(project.skcm), value=T) %>% sort
trg.alleles = c(temp, setdiff(trg.alleles , temp))
trg.crispr.cor = WGCNA::corAndPvalue(project.skcm[trg.alleles,] %>% t, project.skcm[tls.genes,] %>% t)
library(corrplot)
pdf("results/.figs/gdt/cor.trg.tls.pdf", width=8, height=8)
# corrplot(trg.crispr.cor$cor, p.mat = trg.crispr.cor$p, insig = "label_sig",
         # sig.level = c(1E-8, 1E-3, .01), pch.cex = .9, pch.col = "white")
corrplot(trg.crispr.cor$cor[,1:7], method="square")
corrplot(trg.crispr.cor$cor[,7:12], method="square")

dev.off()

trg.crispr.cor = WGCNA::corAndPvalue(x=project.skcm[trg.alleles,] %>% t, 
	y=(project.skcm[setdiff(rownames(project.skcm), trg.alleles),] %>% t))
# avi.dt = data.table(
# 	cor=trg.crispr.cor$cor["TRDV2",], 
# 	p=trg.crispr.cor$p["TRDV2",],
# 	gene = colnames(trg.crispr.cor$cor)
# 	)
# avi.dt = avi.dt[!is.na(p)]



avi.dt = trg.crispr.cor$cor[c("TRDV2", "TRDV1", "TRDV3", "TRBV2"),] %>% t  %>% data.table %>%
	.[,gene:= colnames(trg.crispr.cor$cor)] %>%
	.[!(is.na(TRDV1) | is.na(TRDV2))]

# aa = find.outliers(trg.crispr.cor$cor["TRDV1",], trg.crispr.cor$cor["TRDV2",], limits=c(100,1/100))
aa = fit.loess(avi.dt$TRDV1, avi.dt$TRDV2)
inx.pos = (avi.dt$TRDV2 - aa$fitted) %>% order(decreasing=T) %>% 
.[1:40]
inx.neg = (aa$fitted- avi.dt$TRDV2 ) %>% order(decreasing=T) %>% 
.[1:30]
# avi.dt.sub = avi.dt[inx]
avi.dt.sub = avi.dt[unique(c(
	order(avi.dt$TRDV1,decreasing=F)[1:30],
	order(avi.dt$TRDV2,decreasing=F)[1:30],
	order(avi.dt$TRDV1,decreasing=T)[1:40],
	order(avi.dt$TRDV2,decreasing=T)[1:40], 
	inx.pos,
	inx.neg
	))]
p.sc.volcano = ggplot(data=avi.dt, aes(x=TRDV1, y= TRDV2))  + 
geom_point(alpha=0.5) +
theme_minimal(base_size = 12) + theme(legend.position = "bottom") +
geom_text_repel(
	data = avi.dt.sub,
	 aes(x = TRDV1, y = TRDV2, label = gene),    
	 size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
    )  

ggsave(file="results/.figs/gdt/gene.cor.trdv2.trdv1.pdf", p.sc.volcano, width=16, height=10)

avi.sel=avi.dt[TRDV1>0.5 & TRDV2 > 0.4]
p = get.enrich(genes=avi.sel$gene, universe=avi.dt$gene)
ggsave(file="results/.figs/gdt/pathway.cor.trdv2.pos.trdv1.pos.pdf",p)

try({
p = get.enrich(genes=avi.dt[inx.pos]$gene, universe=avi.dt$gene)
ggsave(file="results/.figs/gdt/pathway.cor.trdv2.pos.trdv1.neg.pdf",p)
})

try({
p = get.enrich(genes=avi.dt[inx.neg]$gene, universe=avi.dt$gene)
ggsave(file="results/.figs/gdt/pathway.cor.trdv2.neg.trdv1.pos.pdf",p)
})



avi.dt = trg.crispr.cor$cor[c("TRDV2", "TRDV1", "TRDV3", "TRBV2"),] %>% t  %>% data.table %>%
	.[,gene:= colnames(trg.crispr.cor$cor)] %>%
	.[!(is.na(TRDV2) | is.na(TRDV2))]
aa = fit.loess(avi.dt$TRDV2, avi.dt$TRBV2)
inx.pos = (avi.dt$TRBV2 - aa$fitted) %>% order(decreasing=T) %>% 
.[1:70]
inx.neg = (aa$fitted- avi.dt$TRBV2 ) %>% order(decreasing=T) %>% 
.[1:70]
# avi.dt.sub = avi.dt[inx]
avi.dt.sub = avi.dt[unique(c(
	order(avi.dt$TRDV2,decreasing=F)[1:30],
	order(avi.dt$TRBV2,decreasing=F)[1:30],
	order(avi.dt$TRDV2,decreasing=T)[1:40],
	order(avi.dt$TRBV2,decreasing=T)[1:40], 
	inx.pos,
	inx.neg
	))]
p.sc.volcano = ggplot(data=avi.dt, aes(x=TRDV2, y= TRBV2))  + 
geom_point(alpha=0.5) +
theme_minimal(base_size = 12) + theme(legend.position = "bottom") +
geom_text_repel(
	data = avi.dt.sub,
	 aes(x = TRDV2, y = TRBV2, label = gene),    
	 size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
    )  

ggsave(file="results/.figs/gdt/gene.cor.trbv2.trdv2.pdf", p.sc.volcano, width=16, height=10)

avi.sel=avi.dt[intersect(
	order(avi.dt$TRBV2,decreasing=T)[1:200], 
	order(avi.dt$TRDV2,decreasing=T)[1:200] )]
p = get.enrich(genes=avi.sel$gene, universe=avi.dt$gene)
ggsave(file="results/.figs/gdt/pathway.cor.trbv2.pos.trdv2.pos.pdf",p)

try({
p = get.enrich(genes=avi.dt[inx.pos]$gene, universe=avi.dt$gene)
ggsave(file="results/.figs/gdt/pathway.cor.trbv2.pos.trdv2.neg.pdf",p)
})

try({
p = get.enrich(genes=avi.dt[inx.neg]$gene, universe=avi.dt$gene)
ggsave(file="results/.figs/gdt/pathway.cor.trbv2.neg.trdv2.pos.pdf",p)
})

avi.sel=avi.dt[intersect(
	order(avi.dt$TRBV2,decreasing=F)[1:200], 
	order(avi.dt$TRDV2,decreasing=F)[1:200] )]
p = get.enrich(genes=avi.sel$gene, universe=avi.dt$gene)
ggsave(file="results/.figs/gdt/pathway.cor.trbv2.neg.trdv2.neg.pdf",p)
```

## correlate with singler annotation
```{r}
cor.out = lapply(singleR.dataset, function(ii) calculate.singleR.corelation(mat=project.skcm, ref.data=ii)) 

skcm.cor.out = do.call(rbind,cor.out)
cell.types =  rownames(skcm.cor.out) %>% unique
skcm.cor.unique = sapply( cell.types, function(tt){
	inx = which(rownames(skcm.cor.out) == tt)
	if(length(inx) > 1){
		out = skcm.cor.out[inx,] %>% colMeans(na.rm=T)
	}else{
		out = skcm.cor.out[inx,]
	}
	out
} ) %>% t

trg.singler.cor = WGCNA::corAndPvalue(x=project.skcm[trg.alleles,] %>% t, 
	y=(skcm.cor.unique %>% t))

avi.dt = trg.singler.cor$cor[c("TRDV2", "TRDV1", "TRDV3"),] %>% t  %>% data.table %>%
	.[,gene:= colnames(trg.singler.cor$cor)] %>%
	.[!(is.na(TRDV1) | is.na(TRDV2))]

aa = fit.loess(avi.dt$TRDV1, avi.dt$TRDV2)
inx.pos = (avi.dt$TRDV2 - aa$fitted) %>% order(decreasing=T) %>% 
.[1:40]
inx.neg = (aa$fitted- avi.dt$TRDV2 ) %>% order(decreasing=T) %>% 
.[1:30]
# avi.dt.sub = avi.dt[inx]
avi.dt.sub = avi.dt[unique(c(
	order(avi.dt$TRDV1,decreasing=F)[1:30],
	order(avi.dt$TRDV2,decreasing=F)[1:30],
	order(avi.dt$TRDV1,decreasing=T)[1:40],
	order(avi.dt$TRDV2,decreasing=T)[1:40], 
	inx.pos,
	inx.neg
	))]
p.sc.volcano = ggplot(data=avi.dt, aes(x=TRDV1, y= TRDV2))  + 
geom_point(alpha=0.5) +
theme_minimal(base_size = 12) + theme(legend.position = "bottom") +
geom_text_repel(
	data = avi.dt.sub,
	 aes(x = TRDV1, y = TRDV2, label = gene),    
	 size = 3,
    box.padding = unit(0.35, "lines"),
    point.padding = unit(0.3, "lines")
    )  

ggsave(file="results/.figs/gdt/singler.cor.trdv2.trdv1.pdf", p.sc.volcano, width=16, height=10)


avi.dt[grep("GC", avi.dt$gene)]
grep("Germinal", rownames(skcm.cor.unique), value=T)

lapply(singleR.dataset, 
```
## TLS makers are only important in presenence of gdt cells. 
```{r}
library(coxme)
library(data.table)
library(magrittr)
library(parallel)
library(survival)
library(survminer)

calculate.cox.interaction = function(dtx){
	tryCatch(
	{
		dtx = dtx[!is.na(gene)]
		aa =  coxph(Surv(Survival, Event) ~ gene + cell.type + gene*cell.type, dtx)
		summary(aa)$coefficients[,c(1,5)] %>% t%>% c
	},
	error=  function(e) rep(NA,6)
	)

}

calculate.cox.interaction.tr = function(dtx){
    tryCatch(
    {
        dtx = dtx[!is.na(gene)]
        aa =  coxph(Surv(Survival, Event) ~ gene + trg +  gene*trg , dtx)%>%
      summary 
    	xx = aa$coefficient
        xx["gene:trg",]
    },
    error=  function(e) rep(NA,5),
    warning = function(e) rep(NA,5)
    )
}
tls.cox.out = lapply(tls.genes, function(tls.gene){ 
      dt.skcm[,trg:=project.skcm[tls.gene,]]
      xx = rep(NA,5)
      try({
      	out = coxph(Surv(Survival, Event) ~ trg, data=dt.skcm) %>% summary 
      	xx = out$coefficient["trg",]

      })


tls.interaction.out.skcm = lapply(trg.alleles, function(trg.allele){ 
    dt.skcm[,trg:=project.skcm[trg.allele,]]%>%
      .[,CD8B:=project.skcm["CD8B",]]
    out = mclapply(tls.genes, function(ii) {
        dt.skcm[,gene:=project.skcm[ii,]]%>%
        calculate.cox.interaction.tr
    }, mc.cores = 60) %>% 
    do.call(rbind, .) %>%
    data.table %>% 
    set_colnames(c("estimate", "exp", "se", "z", "P")) %>% 
    .[,gene:=tls.genes]
    out 
}) 
names(tls.interaction.out.skcm) = trg.alleles
tls.interaction.mat = lapply(tls.interaction.out.skcm, function(tt) {
	ifelse(tt$P<0.03, tt$z,0)
} )%>% do.call(rbind, .) %>% 
set_colnames(tls.genes)

sel = apply(tls.interaction.mat,1,function(tt) tt %>% is.na %>% sum) < length(tls.genes)
tls.interaction.mat = tls.interaction.mat[sel,]
pdf("results/.figs/gdt/interaction.trg.pdf", width=16, height=8)
corrplot(tls.interaction.mat, is.corr = FALSE, method = "square")
dev.off()

out = dt.skcm 
avi.dt = out %>%
.[,TRDV2:=project.skcm["TRDV2",]] %>% 
.[,TRDV2:=ifelse(TRDV2> median(TRDV2),"high", "low")] %>% 
.[,CXCL13:=as.numeric(project.skcm["CXCL13",])] %>% 
.[,CXCL13:=ifelse(CXCL13> median(CXCL13),"high", "low")]

fit1 <- survfit( Surv(Survival, Event) ~ CXCL13, data = avi.dt)
p = ggsurvplot_facet( fit1, avi.dt, facet.by = "TRDV2",  palette = "jco", pval = TRUE)
ggsave("results/.figs/gdt/TRDV2.CXCL13.km.pdf")

fit1 <- survfit( Surv(Survival, Event) ~TRDV2 , data = avi.dt)
p = ggsurvplot_facet( fit1, avi.dt, facet.by = "CXCL13",  palette = "jco", pval = TRUE)
ggsave("results/.figs/gdt/CXCL13.TRDV2.km.pdf")


out = dt.skcm 
avi.dt = out %>%
.[,TRDV1:=project.skcm["TRDV1",]] %>% 
.[,TRDV1:=ifelse(TRDV1> median(TRDV1),"TRDV1-high", "TRDV1-low")] %>% 
.[,MS4A1:=as.numeric(project.skcm["MS4A1",])] %>% 
.[,MS4A1:=ifelse(MS4A1> median(MS4A1),"MS4A1-high", "MSV4A1-low")]

fit1 <- survfit( Surv(Survival, Event) ~ MS4A1, data = avi.dt)
p = ggsurvplot_facet( fit1, avi.dt, facet.by = "TRDV1",  palette = "jco", pval = TRUE)
ggsave("results/.figs/gdt/TRDV1.MS4A1.km.pdf")

fit1 <- survfit( Surv(Survival, Event) ~TRDV1 , data = avi.dt)
p = ggsurvplot_facet( fit1, avi.dt, facet.by = "MS4A1",  palette = "jco", pval = TRUE)
ggsave("results/.figs/gdt/MS4A1.TRDV1.km.pdf")

```