## Analysis of murad's dataset

## Enrichment analysis

```{r}
library(tidyr)
crispr = data.table(readxl::read_excel("~/liulab_home/projects/gdt/data/Murad_Screen Hits.xlsx", sheet = 1))

gene.eg.dt = clusterProfiler::bitr(unique(crispr$id), fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
 kk <- clusterProfiler::enrichKEGG(gene = gene.eg.dt$ENTREZID,
                                    organism     = 'hsa',
                                    pvalueCutoff = 0.05)
  p = clusterProfiler::dotplot(kk, showCategory=30)
 dir.create("results/.figs/murad")
 ggsave(filename = "results/.figs/murad/pathway.pdf",p)
 kk <- clusterProfiler::enrichGO(gene = gene.eg.dt$ENTREZID,
 	 OrgDb    = org.Hs.eg.db,
               ont      = "BP",
                                    pvalueCutoff = 0.05)
  p = clusterProfiler::dotplot(kk, showCategory=30)
 ggsave(filename = "results/.figs/murad/go.pathway.pdf",p)
 
```

## TCGA survival analysis 

```{r}

library(coxme)
library(data.table)
library(magrittr)
library(parallel)
library(survival)
library(survminer)

calculate.cox.interaction = function(dtx){
	tryCatch(
	{
		dtx = dtx[!is.na(gene)]
		aa =  coxph(Surv(Survival, Event) ~ gene + cell.type + gene*cell.type, dtx)
		summary(aa)$coefficients[,c(1,5)] %>% t%>% c
	},
	error=  function(e) rep(NA,6)
	)

}

project.exp = readRDS("~/liulab_home/data/tcga/TCGA_gdc/tcgabiolinks/all.tcga.expression.RDS")
library(SummarizedExperiment)

skcm.inx = which(colData(project.exp)$project_id == "TCGA-SKCM")
project.exp.skcm = project.exp[,skcm.inx] 
project.exp.mat = assay(project.exp.skcm) %>%
set_rownames(rowData(project.exp.skcm)$external_gene_name) %>% t %>% 
 scale(., scale=T, center=T) %>% t 
colnames(project.exp.mat) %<>% substring(.,1,16)
project.skcm = project.exp.mat

dt.skcm = colData(project.exp.skcm) %>% as.data.table %>%
.[,.(days_to_death, days_to_last_follow_up, vital_status, dataset=project_id)] %>% 
.[,Event:=ifelse(vital_status=="Dead", 1, 0)]%>% 
.[,Survival:=ifelse(vital_status=="Dead", days_to_death, days_to_last_follow_up)]

crispr.genes = intersect(rownames(project.exp.mat), crispr$id)
crispr.cox.out = lapply(crispr.genes, function(crispr.gene){ 
      dt.skcm[,trg:=project.skcm[crispr.gene,]]
      xx = rep(NA,5)
      try({
      	out = coxph(Surv(Survival, Event) ~ trg, data=dt.skcm) %>% summary 
      	xx = out$coefficient["trg",]

      })
      xx
    }) %>% do.call(rbind, .) %>%
    data.table %>% 
    set_colnames(c("estimate", "exp", "se", "z", "P")) %>% 
    .[,gene:=crispr.genes]

crispr.cox.out[,padj:=p.adjust(P,method="fdr")]

library(EnhancedVolcano)
p = EnhancedVolcano(crispr.cox.out,
	lab = crispr.cox.out$gene,
	x = 'estimate',
	y = 'P',
	pCutoff = 1E-2,
	FCcutoff = .2,
	# pointSize = c(ifelse(crispr.cox.out$padj <  .1, 3, 2)),
	labSize = 4.0,
	legend=c('NS','Estimate','P value',
		'P value & Estimate'),
	legendPosition = 'bottom',
	legendLabSize = 8,
	legendIconSize = 4.0,
	drawConnectors = TRUE,
	widthConnectors = 0.2,
	xlim = c(-.5, .5),
	colAlpha = 0.8,
	colConnectors = 'grey30'
	)
ggsave(file="results/.figs/murad/cox.pdf",p)

gene.order = crispr.cox.out[order(estimate)]$gene
avi.dt = lapply(gene.order, function(crispr.gene) {
	dt.skcm %>% 
	mutate(expr=project.skcm[crispr.gene,],gene=crispr.gene)
}%>% 
mutate(expr.r=ifelse(expr> median(expr), "high", "low"))
	) %>% do.call(rbind,.)

for (start in seq(1,69,16)) {
	end = start +15
	end = ifelse(end>70, 70, end)
	gene.order.curr = gene.order[seq(start,end)]
	avi.curr = avi.dt[gene %in% gene.order.curr]
	avi.curr$gene %<>% factor(.,levels=gene.order.curr)
	fit1 <- survfit( Surv(Survival, Event) ~ expr.r, data = avi.curr)
	p = ggsurvplot_facet( fit1, avi.curr, facet.by = "gene",  palette = "jco", pval = TRUE)
	ggsave(sprintf("results/.figs/murad/cox.%s.pdf", start), width=10, height=8)
}

## Identify if this genes are enriched for correlation with TRD allele levels. 
temp = c("TRDV2", "TRGV9")
trg.alleles = grep("^TR[GD]V",rownames(project.skcm), value=T) %>% sort
trg.alleles = c(temp, setdiff(trg.alleles , temp))
trg.crispr.cor = WGCNA::corAndPvalue(project.skcm[trg.alleles,] %>% t, project.skcm[gene.order,] %>% t)
library(corrplot)
pdf("results/.figs/murad/cor.trg.pdf", width=16, height=8)
# corrplot(trg.crispr.cor$cor, p.mat = trg.crispr.cor$p, insig = "label_sig",
         # sig.level = c(1E-8, 1E-3, .01), pch.cex = .9, pch.col = "white")
corrplot(trg.crispr.cor$cor, method="square")
dev.off()

calculate.cox.interaction.tr = function(dtx){
    tryCatch(
    {
        dtx = dtx[!is.na(gene)]
        aa =  coxph(Surv(Survival, Event) ~ gene + trg +  gene*trg , dtx)%>%
      summary 
    	xx = aa$coefficient
        xx["gene:trg",]
    },
    error=  function(e) rep(NA,5),
    warning = function(e) rep(NA,5)
    )
}

crispr.interaction.out.skcm = lapply(trg.alleles, function(trg.allele){ 
    dt.skcm[,trg:=project.skcm[trg.allele,]]%>%
      .[,CD8B:=project.skcm["CD8B",]]
    out = mclapply(gene.order, function(ii) {
        dt.skcm[,gene:=project.skcm[ii,]]%>%
        calculate.cox.interaction.tr
    }, mc.cores = 60) %>% 
    do.call(rbind, .) %>%
    data.table %>% 
    set_colnames(c("estimate", "exp", "se", "z", "P")) %>% 
    .[,gene:=gene.order]
    out 
}) 
names(crispr.interaction.out.skcm) = trg.alleles
crispr.interaction.mat = lapply(crispr.interaction.out.skcm, function(tt) {
	ifelse(tt$P<0.03, tt$z,0)
} )%>% do.call(rbind, .) %>% 
set_colnames(gene.order)

sel = apply(crispr.interaction.mat,1,function(tt) tt %>% is.na %>% sum) < length(gene.order)
crispr.interaction.mat = crispr.interaction.mat[sel,]
pdf("results/.figs/murad/interaction.trg.pdf", width=16, height=8)
corrplot(crispr.interaction.mat, is.corr = FALSE, method = "square")
dev.off()

## plotting an example 

out = dt.skcm 
avi.dt = out %>%
.[,TRGV9:=project.skcm["TRGV9",]] %>% 
.[,TRGV9:=ifelse(TRGV9> median(TRGV9),"high", "low")] %>% 
.[,BTN3A1:=as.numeric(project.skcm["BTN3A1",])] %>% 
.[,BTN3A1:=ifelse(BTN3A1> median(BTN3A1),"high", "low")]

fit1 <- survfit( Surv(Survival, Event) ~ BTN3A1, data = avi.dt)
p = ggsurvplot_facet( fit1, avi.dt, facet.by = "TRGV9",  palette = "jco", pval = TRUE)
ggsave("results/.figs/murad/TRGV9.BTN3A1.km.pdf")

fit1 <- survfit( Surv(Survival, Event) ~TRGV9 , data = avi.dt)
p = ggsurvplot_facet( fit1, avi.dt, facet.by = "BTN3A1",  palette = "jco", pval = TRUE)
ggsave("results/.figs/murad/BTN3A1.TRGV9.km.pdf")
```

1) Is it possible to correlate the BTN2A1/3A1 and the mevalonate genes to survival across all the different cancer types? Or is that too laborious/computationally expensive?
2) Is it possible to correlate the same genes to Vg9/Vd2 alleles across all the different cancer types?
For both of these questions, I am partly thinking of Figure 2 of Kai and Shirley's Science paper.

```{r}
mevalonate.genes =  data.table(readxl::read_excel("~/liulab_home/projects/gdt/data/Murad_Screen Hits.xlsx", sheet = 2))
mevalonate.genes = mevalonate.genes$id[seq(2,20)]
crispr.genes = c(mevalonate.genes, "BTN2A1", "BTN3A1")
all.cox =list()
for (cancer.curr in unique(dt.curr$dataset)) {
	cancer.curr.inx = dt.curr$dataset==cancer.curr
	dt.cancer.curr = dt.curr[cancer.curr.inx,]
	project.cancer.curr = project.exp.mat[,cancer.curr.inx]

	crispr.cox.out = lapply(crispr.genes, function(crispr.gene){ 
		dt.cancer.curr[,trg:=project.cancer.curr[crispr.gene,]]
		xx = rep(NA,5)
		try({
			out = coxph(Surv(Survival, Event) ~ trg, data=dt.cancer.curr) %>% summary 
			xx = out$coefficient["trg",]
		})
		xx
	}) %>% do.call(rbind, .) %>%
	data.table %>% 
	set_colnames(c("estimate", "exp", "se", "z", "P")) %>% 
	.[,gene:=crispr.genes]

	crispr.cox.out[,padj:=p.adjust(P,method="fdr")]
	crispr.cox.out$cancer.type = cancer.curr
	all.cox[[cancer.curr]]= crispr.cox.out
}
all.cox.dt = all.cox %>% do.call(rbind,.) %>% as.data.table
all.cox.dt[,plot.value:=ifelse(padj<0.05, z, 0)]
out = dcast(all.cox.dt, gene~cancer.type, value.var="plot.value" ) 
out.mat = out[,-1,with=F] %>% as.matrix %>% 
set_rownames(out$gene)
out.mat = out.mat[crispr.genes,]
pdf("~/liulab_home/projects/gdt/results/.figs/murad/pancancer.interaction.trg.melavolante.pdf", width=16, height=8)
corrplot(t(out.mat), is.corr = FALSE, method = "square")
dev.off()

trd2.cor = list()
trg9.cor = list()
for(cancer.curr in unique(dt.curr$dataset)) {
	cancer.curr.inx = dt.curr$dataset==cancer.curr
	dt.cancer.curr = dt.curr[cancer.curr.inx,]
	project.cancer.curr = project.exp.mat[,cancer.curr.inx]
	trdv29.alleles = c("TRDV2", "TRGV9")
	trg.crispr.cor = WGCNA::corAndPvalue(project.cancer.curr[trdv29.alleles,] %>% t, project.cancer.curr[crispr.genes,] %>% t)
	trd2.cor[[cancer.curr]] = trg.crispr.cor$cor["TRDV2",]
	trg9.cor[[cancer.curr]] = trg.crispr.cor$cor["TRGV9",]
}

pdf("~/liulab_home/projects/gdt/results/.figs/murad/pancancer.cor.trd2..melavolante.pdf", width=16, height=8)
corrplot(do.call(rbind,trd2.cor), method="square")
dev.off()
pdf("~/liulab_home/projects/gdt/results/.figs/murad/pancancer.cor.trg9..melavolante.pdf", width=16, height=8)
corrplot(do.call(rbind,trg9.cor), method="square")
dev.off()
```

## Murad hits all genes
# 1. GSEA analysis 
# 2. MSIG db enrichment. Check melavolnte pathway. 
# 3. 

```{r}

library(tidyr)
crispr = data.table(readxl::read_excel("~/liulab_home/projects/gdt/data/Murad_Screen Hits.xlsx", sheet = 1))

gene.eg.dt = clusterProfiler::bitr(unique(crispr$id), fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
 kk <- clusterProfiler::enrichKEGG(gene = gene.eg.dt$ENTREZID,
                                    organism     = 'hsa',
                                    pvalueCutoff = 0.05)
  p = clusterProfiler::dotplot(kk, showCategory=30)
 dir.create("results/.figs/murad")
 ggsave(filename = "results/.figs/murad/pathway.pdf",p)
 kk <- clusterProfiler::enrichGO(gene = gene.eg.dt$ENTREZID,
 	 OrgDb    = org.Hs.eg.db,
               ont      = "BP",
                                    pvalueCutoff = 0.05)
  p = clusterProfiler::dotplot(kk, showCategory=30)
 ggsave(filename = "results/.figs/murad/go.pathway.pdf",p)
 
```
## TCGA survival analysis 


```
