---
title: "Analyze David Ting data"
output: html_notebook
---


###
For reference we need count, cell type and nUMI
Because we need to useread count, we need only cell annotations of for reference datasets. 
Most appropiate dataset would be Livnat dataset. But there could be problem in the dataset
Another option is to use melanoma datasets from TISCH.

## Get (Avi) Livnat dataset
Check annotations of gamma-delta T-cells, and Melonoma cells 

```{r}
library(Seurat)
icb.integrated = readRDS("~/liulab_home/data/single_cell/all.icb.integrated.rds")
metadata.list = readRDS("/homes6/asahu/liulab_home/data/single_cell//all.icb.integrated.metadata.rds")
all.meta = icb.integrated@meta.data
datasets.num = lapply(rownames(all.meta), function(tt) {
  xx = strsplit(tt,split = "_") %>% unlist
  xx[length(xx)]
}
  ) %>% unlist 

temp = icb.integrated[,which(datasets.num=="13")]#part1.sub50000
p1=DimPlot(temp, label=T, raster=T)+ NoLegend(); p2=  DimPlot(temp, label=T, group.by = "celltype", raster=T) + NoLegend()
ggsave("~/liulab_home/projects/gdt/results/rctd/icb.all.integrated.covid.pdf", plot = p1+p2, width = 15, height = 10)

temp = icb.integrated[,which(datasets.num=="9")]#livnat 
p1=DimPlot(temp, label=T, raster=T)+ NoLegend();
ggsave("~/liulab_home/projects/gdt/results/rctd/icb.all.integrated.livnat.pdf", plot = p1, width = 15, height = 10)

```

Annotate clusters
```{r}
DimPlot(icb.integrated[,icb.integrated$dataset.num==9])
```

label transfer to all skcm datasets

```{r}
library(Seurat)
covid.obj = readRDS("/liulab/asahu/data/covid/scarlettQ/part1.sub50000.Rds")


icb.objs = list.files(path="/liulab/zzeng/Tigger/static/data/scICB/", pattern = "*.rds",recursive = T, full.names = T) %>% grep("SKCM", ., value=T)

nonicb.objs = list.files(path="/liulab/zzeng/Tigger/static/data/scNonICB/", pattern = "*.rds",recursive = T, full.names = T)%>% grep("SKCM", ., value=T)

 covid.obj%<>% NormalizeData %>% 
FindVariableFeatures(selection.method = "vst", 
        nfeatures = 3000, verbose = FALSE) %>%
scaleData %>% 
RunPCA(pc.genes=VariableFeatures(covid.obj@var.genes), num.dim=30)



scRNA.file="/liulab/zzeng/Tigger/static/data/scNonICB//SKCM_GSE72056/SKCM_GSE72056_res.rds"
sco.curr = get.sco.label(icb.objs[1])

p = DimPlot(sco, group.by="celltype", label=T) + NoLegend()
ggsave(sprintf("%s/dimplot_GSE72056.pdf", resultsdir))

all.meta.covid.summ = all.meta.covid[, .N, by=.(celltype,seurat_clusters) ]
all.meta.covid.summ[,total.celltype:=sum(N),by=.(celltype)]
all.meta.covid.summ[,total.seurat_clusters:=sum(N),by=.(seurat_clusters)]
all.meta.covid.summ[,jaccard:=N/(total.celltype+total.seurat_clusters -N),by=.(seurat_clusters)]
all.meta.covid.summ[,max.jaccard:=max(jaccard), by=.(seurat_clusters)]
all.meta.covid.summ2=all.meta.covid.summ[max.jaccard==jaccard]

```

### run rctd

Create reference 
```{r}
library(RCTD)
library(Matrix)

counts = sco@assays$RNA@counts %>% round
# cell_type = sco$predicted.id %>% as.factor
cell_type = sco$celltype %>% as.factor

cell_type.sel = which(table(cell_type) > 25) %>% names 
sel.inx = which(cell_type %in%  cell_type.sel)
cell_type.curr =  cell_type[sel.inx] 
cell_type.curr = factor(cell_type.curr, levels=unique(cell_type.curr))
reference <- Reference(counts=counts[,sel.inx], cell_types=cell_type.curr)

## Examine reference object (optional)
print(dim(reference@counts)) #observe Digital Gene Expression matrix
table(reference@cell_types) #number of occurences for each cell type

## Save RDS object (optional)
refdir = "~/liulab_home/projects//gdt//results//rctd/references"
saveRDS(reference, file.path(refdir,'SKCM_GSE72056.rds'))

```

Read spatial data from David 

```{r}
source("source.R")
counts =fread("/liulab/asahu/projects/gdt/data/ting_07092021/rawCounts.tsv")
meta.data =readxl::read_excel("/liulab/asahu/projects/gdt/data/ting_07092021/Melanoma_20210601_meta.xlsx", sheet = 1)
meta.data$sample=paste0(meta.data$Sample_ID, ".dcc")
count.mat = dt2mat(counts)
count.curr = round(count.mat)


sco.spatial = CreateSeuratObject(counts=count.mat)
sco.spatial = preprocessing.geomx.seurat(sco.spatial,num.dim = 20)

sco.spatial$patients = meta.data$`Slide name`[match(meta.data$sample,colnames(sco.spatial))]
spatial$tumor_or_immune = meta.data$Comments[match(meta.data$sample,colnames(sco.spatial))]

Idents(sco.spatial) = sco.spatial$patients
p=Seurat::DimPlot(sco.spatial)
## using umap as coords
coords.curr = sco.spatial@reductions$umap@cell.embeddings %>% as.data.frame
nUMI <- colSums(count.mat) %>%round# In this case, total counts per pixel is nUMI

### Create SpatialRNA object
puck <- SpatialRNA(coords.curr, count.curr, nUMI)

myRCTD <- create.RCTD(puck, reference, max_cores = 10)
myRCTD <- run.RCTD(myRCTD, doublet_mode = 'full')
saveRDS(file="~/liulab_home/projects//gdt//results//rctd/references/sco.spatial.Rds", sco.spatial)

```

Analyze result
```{r}
results <- myRCTD@results
# normalize the cell type proportions to sum to 1.
norm_weights = sweep(results$weights, 1, rowSums(results$weights), '/') 
cell_type_names <- myRCTD@cell_type_info$info[[2]] #list of cell type names
spatialRNA <- myRCTD@spatialRNA
resultsdir <- '~/liulab_home/projects//gdt//results//rctd/RCTD_Plots' ## you may change this to a more accessible directory on your computer.
dir.create(resultsdir)
```
---
Below I perform preliminary results
1. if tumor has higher malignant weights
2. plot heatmap 


```{r}
avi.dt = data.table(tumor=sco.spatial$tumor_or_immune %>% as.factor, mal.weights=norm_weights[,"Malignant"])
avi.dt = avi.dt[!is.na(tumor)]
library(ggpubr)
p = ggboxplot(avi.dt, x="tumor", y = "mal.weights", add = "dotplot")
filename = sprintf("%s/tumor_weights.pdf", resultsdir)
ggsave(file=filename, p)
library(corrplot)
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF",  "#FFFFFF", "#77AADD", "#4477AA") %>% rev)
ord=sco.spatial$tumor_or_immune %>% order(na.last="FALSE")
label = sco.spatial$tumor_or_immune[ord]
label[1] = "control"
weights.ord = as.matrix(norm_weights)[ord,] %>% t %>% 
set_colnames(label)
pdf(sprintf("%s/weights_matrix.pdf", resultsdir), width=16, height=6)
corrplot(weights.ord, method="color", col=col(200),
         # type="upper",
         # order="hclust",
         is.corr = F,
         # addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=90, #Text label color and rotation
         tl.cex = 0.8,
         # Combine with significance
         # p.mat = p.mat, sig.level = 0.01, insig = "blank",
         # hide correlation coefficient on the principal diagonal
         diag=T
         )
dev.off()
```


plot curated annotation 
```r
  resultsdir2 <- '~/liulab_home/projects//gdt//results//rctd/RCTD_Plots' ## you may change this to a more accessible directory on your computer.
puck.immune = get.spatial.obj(sco.spatial[,sco.spatial$tumor_or_immune=="Immune"])
# out = get.RTCD.results(sco=sco, annotation="assign.ident", resultsdir=sprintf("%s/tisch", resultsdir2))
# out.level1_annot = get.RTCD.results(sco=sco, annotation="assign.level1_anno", resultsdir=sprintf("%s/level1_anno", resultsdir2))
out.assign.curated = get.RTCD.results(sco=sco, rctd.spatial= puck, annotation="assign.curated", resultsdir=sprintf("%s/assign.curated", resultsdir2))
out.predicted.id = get.RTCD.results(sco=sco, rctd.spatial= puck, annotation="predicted.id", resultsdir=sprintf("%s/predicted.id", resultsdir2))
```



```r
run.reference = function(resultsdir2, reffile, immune.only=F){
  source("source.R")
  dir.create(resultsdir2)
  sco.spatial.curr = sco.spatial
  if(immune.only) sco.spatial.curr = sco.spatial[,sco.spatial$tumor_or_immune=="Immune"]
  puck.immune = get.spatial.obj(sco.spatial.curr)

  sco = get.sco.label(reffile)

try({
  out.assign.paper = get.RTCD.results(
    sco=sco, rctd.spatial= puck, annotation="assign.paper", resultsdir=sprintf("%s/assign.paper", resultsdir2), tumor_or_immune = sco.spatial.curr$tumor_or_immune)
})
try({
  out.assign.curated = get.RTCD.results(sco=sco, rctd.spatial= puck, annotation="assign.curated", resultsdir=sprintf("%s/assign.curated", resultsdir2), tumor_or_immune = sco.spatial.curr$tumor_or_immune)
})
try({out.predicted.id = get.RTCD.results(sco=sco, rctd.spatial= puck, annotation="predicted.id", resultsdir=sprintf("%s/predicted.id", resultsdir2), tumor_or_immune = sco.spatial.curr$tumor_or_immune)
})

sco
}

reffile = "/liulab/zzeng/Tigger/static/data/scNonICB//SKCM_GSE123139/SKCM_GSE123139_res.rds"
resultsdir2 <- '~/liulab_home/projects//gdt//results//rctd/SKCM_GSE123139' 
sco.SKCM_GSE123139= run.reference(resultsdir2=resultsdir2,reffile, immune.only=T)
```